---
title: "SignBase: Maps for Aurignacian Sites"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
R version: R version 3.4.4 (2018-03-15) -- "Someone to Lean On"
Linux version: Ubuntu 18.04.3 LTS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

# Load libraries
Load the following packages. If these are not yet installed use install.packages("") to install them.
```{r}
library(ggmap)
library(plyr)
library(gridExtra)
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(scatterplot3d)
library(googleway)
library(rgl)
library(plotly)
library(grid)
```

# Load data
Run this code to load the file with data on Aurignacian objects.
```{r}
objects <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-04-25.csv")
nrow(objects)
# Select Aurignacian objects from Swabian Jura Caves
objects <- objects[objects$archaeological_unit == "Aurignacien" &
                     objects$site_name == "Vogelherd" |
                     objects$site_name == "Hohle Fels" |
                     objects$site_name == "Geissenklösterle" |
                     objects$site_name == "Hohlenstein-Stadel", ]
nrow(objects)
```
# Preprocessing

Create data frame with object counts per site.
```{r}
sites <- count(objects, site_name)
colnames(sites) <- c("site_name", "number_of_objects")
```

# Collapse Factors
Collapse some factors of meta information which are uncommon and/or related. 
```{r}
# material
table(objects$material)
# object_type
table(objects$object_type)

# material: collapse manually
objects$material <- fct_collapse(objects$material,
             ivory = "ivory",
             antler = "antler",
             bone = "bone",
             other_level = "other")

# object_type: collapse manually
objects$object_type <- fct_collapse(objects$object_type,
             figurine_anthropomorph = c("figurine anthropomorph",
                          "plaquette"), 
             # note: the "plaquette" is the so-called Adorant with an anthropomorph figure on it.
             figurine_zoomorph = c("figurine zoomorph"),
             tool = c("awl", "bâton percé",
                      "rod/bâton", "spatula/lissoir"),
             tube_flute = c("flute", "tube", "possible flute", "tube/flute"),
             personal_ornament = c("personal ornament"),
             other_level = "other")
```

## ggmap with OBJECT locations
Set API key. 
```{r}
register_stadiamaps(key = "c56e7665-ab34-4608-8730-428e35d2766e", write = FALSE)
register_google(key = "AIzaSyDVDHFLK0KnDoxuRfEauKSzvU1Udxu92RY")
```

Map of Germany with Swabian Jura indication. 
```{r}
# Germany
germany.map.background <- get_map(location = "Germany", maptype = "stamen_toner_background", zoom = 6)
germany.map <- ggmap(germany.map.background) +
  geom_rect(aes(xmin = 9.7, xmax = 10.25, ymin = 48.35, ymax = 48.7), color = "red", fill = NA) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
germany.map
```

Plot a map of the Swabian Jura sites including Lone Valley and Ach Valley with each object and its location, and the densities of objects. Note that the object locations need to be jittered, otherwise they would just be overplotted for all objects of the same site.
```{r, fig.width = 5, fig.height = 5}
# Achtal and Lonetal, get map
map <- get_stadiamap(bbox = c(left = 9.7, bottom = 48.35, right = 10.25, top = 48.7), 
                         zoom = 12, maptype = c("stamen_terrain"), crop = TRUE, messaging = FALSE,
                         urlonly = FALSE, force = FALSE,
                         where = tempdir())
ggmap(map)

# Achtal + Lonetal Plot
object.density.map <- ggmap(map) +
  stat_density2d(data = objects,
                 aes(x = longitude, y = latitude, fill = ..level..),
                 alpha = 0.1, size = 0.1,
                 geom = "polygon") +
  geom_density2d(data = objects, 
                 aes(x = longitude, y = latitude), size = 0.3, colour = "white") +
  scale_fill_gradient(low = "yellow", high= "red") +
  geom_point(data = objects,
             aes(jitter(longitude, amount = 0.005), jitter(latitude, amount = 0.005)), 
             size = 0.2)  +
  theme(legend.position = "none") +
  ggtitle("Swabian Jura") +
  annotate("segment", x = 10.18, xend = 10, y = 48.559, yend = 48.559, colour = "black") +
  annotate("text", x = 10.05, y = 48.57, label = "Lone Valley", colour = "black") +
  annotate("segment", x = 9.77, xend = 9.85, y = 48.385, yend = 48.385, colour = "black") +
  annotate("text", x = 9.85, y = 48.395, label = "Ach Valley", colour = "black") +
  annotate("text", x = 10, y = 48.41, label = "Ulm", colour = "black") +
  annotate("segment", x = 10.18, xend = 9.77, y = 48.559, yend = 48.385, colour = "black", linetype = 3) +
   annotate("text", x = 10.02, y = 48.47, label = "c. 35 km", colour = "black")
swabianJura.map <- object.density.map +
  # add map of Germany as inset
  inset(ggplotGrob(germany.map), xmin = 9.72, xmax = 9.98, ymin = 48.5, ymax = 48.69)
```

Subset object data frame according to cave site.
```{r}
# separate datasets per cave site
objects.gkl <- objects[objects$site_name == "Geissenklösterle", ]
objects.hfc <- objects[objects$site_name == "Hohle Fels", ]
objects.vhc <- objects[objects$site_name == "Vogelherd", ]
objects.hss <- objects[objects$site_name == "Hohlenstein-Stadel", ]
```

# Maps
Maps with objects for Achtal and Lonetal.
```{r}
# Achtal, get map
map.achtal <- get_map(location = c(lon = 9.76473, lat = 48.39075), 
                         zoom = 13, maptype = "satellite", source = "google")
ggmap(map.achtal)

# Achtal Plot
object.map.achtal <- ggmap(map.achtal) +
  # Geissenkösterle
  geom_text_repel(data = objects.gkl, 
                  aes(x = longitude,
                      y = latitude,
                      label = " "),
                  color = "white",
                  direction = 'y',
                  nudge_x = 0.041,
                  segment.size = 0.2,
                  size = 2,
                  max.overlaps = Inf) +
  # Hohle Fels
  geom_text_repel(data = objects.hfc, 
                  aes(x = longitude,
                      y = latitude,
                      label = " "),
                  color = "white",
                  direction = 'y',
                  nudge_x = -0.04,
                  segment.size = 0.2,
                  size = 1.8,
                  max.overlaps = Inf) +
  ggtitle("Ach Valley") +
  theme(legend.position = c(.75, .2)) +
  annotate("segment", x = 9.772222, y = 48.398333, xend = 9.754103, yend = 48.379117, color = "white",
               linetype = 3) +
  annotate("text", x = 9.768, y = 48.389, label = "2.5 km", colour = "white") +
  annotate("segment", x = 9.772222, xend = 9.772222, y = 48.398333, yend = 48.413, colour = "white") +
  annotate("text", x = 9.772222, y = 48.415, label = "Geissenklösterle", colour = "white") +
  annotate("segment", x = 9.754103, xend = 9.754103, y = 48.379117, yend = 48.37, colour = "white") +
  annotate("text", x = 9.754103, y = 48.368, label = "Hohle Fels", colour = "white") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
object.map.achtal
```

Plot a map of the Lonetal.
```{r}
# Lonetal, get map
map.lonetal <- get_map(location = c(lon = 10.185, lat = 48.5546), 
                         zoom = 15, maptype = "satellite", source = "google")
ggmap(map.lonetal)

# Lonetal plot
object.map.lonetal <- ggmap(map.lonetal) +
   # Vogelherd
  geom_text_repel(data = objects.vhc, 
                  aes(x = longitude,
                      y = latitude,
                      label = " "),
                  color = "white",
                  direction = 'y',
                  nudge_x = 0.004,
                  segment.size = 0.2,
                  size = 2,
                  max.overlaps = Inf) +
  # Hohlenstein-Stadel
  geom_text_repel(data = objects.hss, 
                  aes(x = longitude,
                      y = latitude,
                      label = sign_coding),
                  color = "white",
                  direction = 'y',
                  nudge_x = 0.006,
                  nudge_y = -0.002,
                  segment.size = 0.2,
                  size = 2,
                  max.overlaps = Inf) +
  geom_segment(aes(x = 10.194, y = 48.559, xend = 10.17255, yend = 48.54975), color = "white",
               linetype = 3) +
  theme(legend.position = "none") +
  ggtitle("Lone Valley") +
  annotate("text", x = 10.185, y = 48.554, label = "1.9 km", colour = "white") +
  annotate("segment", x = 10.194, xend = 10.19, y = 48.559, yend = 48.559, colour = "white") +
  annotate("text", x = 10.19, y = 48.56, label = "Vogelherd", colour = "white") +
  annotate("segment", x = 10.17255, xend = 10.175, y = 48.54975, yend = 48.54975, colour = "white") +
  annotate("text", x = 10.18, y = 48.550, label = "Hohlenstein-Stadel", colour = "white") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
object.map.lonetal
```

Add plots with codings by cave site.
```{r, fig.width = 5, fig.height = 5}
set.seed(09012020)
# Geissenklösterle
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.gkl.sample <- objects.gkl[objects.gkl$object_type != "other", ]
# sample subset of objects randomly (otherwise the plot is overcrowded)
objects.gkl.sample <- sample_n(objects.gkl.sample, 25)
# Plot
coding.plot.gkl <- ggplot(objects.gkl.sample, 
                          aes(x = rep(0, nrow(objects.gkl.sample)), y = object_id)) +
  geom_text(aes(label = sign_coding), hjust = 0, nudge_x = 0.3 , size = 2.5) +
  geom_text(aes(label = object_type, color = object_type), hjust = 0, nudge_x = 0 , size = 2.5) +
  xlim(0, 1) +
  theme_classic2() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position = "none") +
  ggtitle(paste("Geissenklösterle, n = ", nrow(objects.gkl) , sep = ""))
coding.plot.gkl

# Hohle Fels
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.hfc.sample <- objects.hfc[objects.hfc$object_type != "other", ]
# sample subset of objects randomly (otherwise the plot is overcrowded)
objects.hfc.sample <- sample_n(objects.hfc.sample, 25)
# Plot
coding.plot.hfc <- ggplot(objects.hfc.sample, 
                          aes(x = rep(0, nrow(objects.hfc.sample)), y = object_id)) +
  geom_text(aes(label = sign_coding), hjust = 0, nudge_x = 0.3 , size = 2.5) +
  geom_text(aes(label = object_type, color = object_type), hjust = 0, nudge_x = 0 , size = 2.5) +
  xlim(0, 1) +
  theme_classic2() + 
  #scale_y_discrete(position = "right") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position = "none") +
  ggtitle(paste("Hohle Fels, n = ", nrow(objects.hfc) , sep = ""))
coding.plot.hfc

# Vogelherd
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.vhc.sample <- objects.vhc[objects.vhc$object_type != "other", ]
# sample subset objects randomly (otherwise the plot is overcrowded)
objects.vhc.sample <- sample_n(objects.vhc.sample, 25)
# Plot
coding.plot.vhc <- ggplot(objects.vhc.sample, 
                          aes(x = rep(0, nrow(objects.vhc.sample)), y = object_id)) +
  geom_text(aes(label = sign_coding), hjust = 0, nudge_x = 0.3 , size = 2.5) +
  geom_text(aes(label = object_type, color = object_type), hjust = 0, nudge_x = 0 , size = 2.5) +
  # geom_point(size = 1) +
  xlim(0, 1) +
  theme_classic2() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position = "none") +
  ggtitle(paste("Vogelherd, n = ", nrow(objects.vhc) , sep = ""))
coding.plot.vhc
```

Combine plots.
```{r, fig.width = 15, fig.height = 10}
combined.plot <- ggarrange(swabianJura.map, object.map.lonetal, coding.plot.vhc, 
                           coding.plot.hfc, object.map.achtal, coding.plot.gkl, ncol = 3, nrow = 2,
                           labels = "AUTO")
combined.plot
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/maps/maps_combined.pdf", 
       combined.plot, dpi = 500, width = 17, height = 10, device = cairo_pdf)
```

# 3D Plot
```{r}
# set google API key
# set_key("AIzaSyDVDHFLK0KnDoxuRfEauKSzvU1Udxu92RY")
# 
# ## elevation data for the Achtal
# achtal.df <- data.frame(lat = sample(9740:9790, 50, replace = F)/1000,
#                  lon = sample(48370:48420, 50, replace = F)/1000)
# 
# achtal.df <- google_elevation(df_locations = achtal.df,
#                        location_type = "individual",
#                        samples = 100,
#                        simplify = TRUE)
# 
# df <- data.frame(achtal.df$results$location$lng,
#                    achtal.df$results$location$lat,
#                    achtal.df$results$elevation)
# 
# colnames(df) <- c("longitude", "latitude", "elevation")
# 
# s3d <- scatterplot3d(x = achtal.df$results$location$lng, 
#                      y = achtal.df$results$location$lat, 
#                      z = achtal.df$results$elevation, 
#                      type = "p", 
#                      pch = 1, 
#                      highlight.3d = TRUE, 
#                      cex.symbols = 0.5, 
#                      box = FALSE,
#                      grid = FALSE,
#                      xlab = "Longitude",
#                      ylab = "Latitude",
#                      zlab = "Elevation")
# print(s3d)
# 
# achtal.plot <- plot_ly(x = achtal.df$results$location$lng, 
#                      y = achtal.df$results$location$lat, 
#                      z = achtal.df$results$elevation,
#                      type = "mesh3d", 
#                      #color = achtal.df$results$elevation,
#                      #colors = "YlOrRd",
#                      height = 1000)
# 
# achtal.matrix <- acast(df, longitude~latitude, value.var="elevation")
# #spread(df, latitude, elevation)
# 
# persp3D(x = df$longitude, y = df$latitude, z = achtal.matrix, main = "volcano", clab = c("height", "m"),
# breaks = seq(0,1000, by = 100))
```


