---
title: "SignBase: Maps for Aurignacian Sites"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
R version: R version 3.4.4 (2018-03-15) -- "Someone to Lean On"
Linux version: Ubuntu 18.04.3 LTS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

# Load libraries
Load the following packages. If these are not yet installed use install.packages("") to install them.
```{r}
library(ggmap)
library(plyr)
library(gridExtra)
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(scatterplot3d)
library(googleway)
library(rgl)
library(plotly)
```

# Load data
Run this code to load the file with data on Aurignacian objects.
```{r}
objects <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-04-22.csv")
nrow(objects)
# Select Aurignacian objects from Swabian Jura Caves
objects <- objects[objects$archaeological_unit == "Aurignacien" &
                     objects$site_name == "Vogelherd" |
                     objects$site_name == "Hohle Fels" |
                     objects$site_name == "Geissenklösterle" |
                     objects$site_name == "Hohlenstein-Stadel", ]
nrow(objects)
```
# Preprocessing

Create data frame with object counts per site.
```{r}
sites <- count(objects, site_name)
colnames(sites) <- c("site_name", "number_of_objects")
```

# Collapse Factors
Collapse some factors of meta information which are uncommon and/or related. 
```{r}
# material
table(objects$material)
# object_type
table(objects$object_type)

# material: collapse manually
objects$material <- fct_collapse(objects$material,
             ivory = "ivory",
             antler = "antler",
             bone = "bone",
             other_level = "other")

# object_type: collapse manually
objects$object_type <- fct_collapse(objects$object_type,
             figurine_anthropomorph = c("figurine anthropomorph",
                          "plaquette"), 
             # note: the "plaquette" is the so-called Adorant with an anthropomorph figure on it.
             figurine_zoomorph = c("figurine zoomorph"),
             tool = c("awl", "bâton percé",
                      "rod/bâton", "spatula/lissoir"),
             tube_flute = c("flute", "tube", "possible flute", "tube/flute"),
             personal_ornament = c("personal ornament"),
             other_level = "other")
```

## ggmap with OBJECT locations
Set API key. 
```{r}
register_stadiamaps(key = "c56e7665-ab34-4608-8730-428e35d2766e", write = FALSE)
register_google(key = "AIzaSyDVDHFLK0KnDoxuRfEauKSzvU1Udxu92RY")
```

Load stamenmap of Swabian Jura locations. 
```{r}
# Achtal and Lonetal
map <- get_stadiamap(bbox = c(left = 9.7, bottom = 48.35, right = 10.25, top = 48.6), 
                         zoom = 12, maptype = c("stamen_terrain"), crop = TRUE, messaging = FALSE,
                         urlonly = FALSE, force = FALSE,
                         where = tempdir())
ggmap(map)

# Achtal
map.achtal <- get_map(location = c(lon = 9.76473, lat = 48.39075), 
                         zoom = 14, maptype = "satellite", source = "google")
ggmap(map.achtal)

# Lonetal
map.lonetal <- get_map(location = c(lon = 10.185, lat = 48.5546), 
                         zoom = 15, maptype = "satellite", source = "google")
ggmap(map.lonetal)

```

Plot a map of Europe with each object and its location, and the densities of objects. Note that the object locations need to be jittered, otherwise they would just be overplotted for all objects of the same site.
```{r}
# Achtal + Lonetal
object.density.map <- ggmap(map) +
  stat_density2d(data = objects,
                 aes(x = longitude, y = latitude, fill = ..level..),
                 alpha = 0.1, size = 0.1,
                 geom = "polygon") +
  geom_density2d(data = objects, 
                 aes(x = longitude, y = latitude), size = 0.3, colour = "white") +
  scale_fill_gradient(low = "yellow", high= "red") +
  geom_point(data = objects,
             aes(jitter(longitude, amount = 0.005), jitter(latitude, amount = 0.005)), 
             size = 0.2)  +
  theme(legend.position = "none") +
  ggtitle("Swabian Jura") +
  annotate("segment", x = 10.18, xend = 10, y = 48.559, yend = 48.559, colour = "black") +
  annotate("text", x = 10, y = 48.57, label = "Lone Valley", colour = "black") +
  annotate("segment", x = 9.77, xend = 9.85, y = 48.385, yend = 48.385, colour = "black") +
  annotate("text", x = 9.85, y = 48.395, label = "Ach Valley", colour = "black") +
  annotate("text", x = 10, y = 48.41, label = "Ulm", colour = "black") +
  annotate("segment", x = 10.18, xend = 9.77, y = 48.559, yend = 48.385, colour = "black", linetype = 3) +
   annotate("text", x = 10.02, y = 48.47, label = "c. 35 km", colour = "black")
object.density.map
```

Plot a map of the Achtal.
```{r}
# separate datasets per cave site
objects.gkl <- objects[objects$site_name == "Geissenklösterle", ]
# Order by a given column (e.g. object_type)
objects.gkl.ordered <- objects.gkl[order(objects.gkl$object_type), ]

# Achtal
object.map.achtal <- ggmap(map.achtal) +
  # stat_density2d(data = objects,
  #                aes(x = longitude, y = latitude, fill = ..level..),
  #                alpha = 0.3, size = 0.1,
  #                geom = "polygon") +
  # geom_density2d(data = objects, 
  #               aes(x = longitude, y = latitude), size = 0.3, colour = "white") +
  #scale_fill_gradient(low = "yellow", high= "red") +
  geom_point(data = objects,
             aes(jitter(longitude, amount = 0.005), jitter(latitude, amount = 0.005),
                 color = object_type),
             size = 2)  +
  ggtitle("Ach Valley") +
  theme(legend.position = c(.75, .2)) +
  annotate("segment", x = 9.772222, y = 48.398333, xend = 9.754103, yend = 48.379117, color = "white",
               linetype = 3) +
  annotate("text", x = 9.768, y = 48.389, label = "2.5 km", colour = "white") +
  annotate("segment", x = 9.772222, xend = 9.75, y = 48.398333, yend = 48.4, colour = "white") +
  annotate("text", x = 9.75, y = 48.401, label = "Geissenklösterle", colour = "white") +
  annotate("segment", x = 9.754103, xend = 9.75, y = 48.379117, yend = 48.39, colour = "white") +
  annotate("text", x = 9.75, y = 48.391, label = "Hohle Fels", colour = "white")
object.map.achtal
```

Plot a map of the Lonetal.
```{r}
# separate datasets per cave site
objects.vhc <- objects[objects$site_name == "Vogelherd", ]
# Order by a given column (e.g. object_type)
objects.vhc.ordered <- objects.vhc[order(objects.vhc$object_type), ]

objects.hss <- objects[objects$site_name == "Hohlenstein-Stadel", ]
# Order by a given column (e.g. object_type)
objects.hss.ordered <- objects.hss[order(objects.hss$object_type), ]

# Lonetal
object.map.lonetal <- ggmap(map.lonetal) +
  geom_point(data = objects.vhc,
             aes(jitter(longitude, amount = 0.003), jitter(latitude, amount = 0.003),
                 color = object_type),
             size = 2) +
  geom_point(data = objects.hss,
             aes(jitter(longitude, amount = 0.001), jitter(latitude, amount = 0.001),
                 color = object_type),
             size = 2) +
  geom_segment(aes(x = 10.194, y = 48.559, xend = 10.17255, yend = 48.54975), color = "white",
               linetype = 3) +
  theme(legend.position = "none") +
  ggtitle("Lone Valley") +
  annotate("text", x = 10.185, y = 48.554, label = "1.9 km", colour = "white") +
  annotate("segment", x = 10.194, xend = 10.182, y = 48.559, yend = 48.560, colour = "white") +
  annotate("text", x = 10.182, y = 48.5605, label = "Vogelherd", colour = "white") +
  annotate("segment", x = 10.17255, xend = 10.18, y = 48.54975, yend = 48.550, colour = "white") +
  annotate("text", x = 10.185, y = 48.550, label = "Hohlenstein-Stadel", colour = "white") 
object.map.lonetal
```


## Put both maps together in one panel plot
```{r, fig.width = 18, fig.height = 7}
maps.combined <- ggarrange(object.density.map, object.map.achtal, object.map.lonetal, ncol = 3, widths = c(1.5,1,1))
maps.combined
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/maps/maps_combined.pdf", 
       maps.combined, dpi = 300, width = 18, height = 7, device = cairo_pdf)
```

# 3D Plot
```{r}
# set google API key
# set_key("AIzaSyDVDHFLK0KnDoxuRfEauKSzvU1Udxu92RY")
# 
# ## elevation data for the Achtal
# achtal.df <- data.frame(lat = sample(9740:9790, 50, replace = F)/1000,
#                  lon = sample(48370:48420, 50, replace = F)/1000)
# 
# achtal.df <- google_elevation(df_locations = achtal.df,
#                        location_type = "individual",
#                        samples = 100,
#                        simplify = TRUE)
# 
# df <- data.frame(achtal.df$results$location$lng,
#                    achtal.df$results$location$lat,
#                    achtal.df$results$elevation)
# 
# colnames(df) <- c("longitude", "latitude", "elevation")
# 
# s3d <- scatterplot3d(x = achtal.df$results$location$lng, 
#                      y = achtal.df$results$location$lat, 
#                      z = achtal.df$results$elevation, 
#                      type = "p", 
#                      pch = 1, 
#                      highlight.3d = TRUE, 
#                      cex.symbols = 0.5, 
#                      box = FALSE,
#                      grid = FALSE,
#                      xlab = "Longitude",
#                      ylab = "Latitude",
#                      zlab = "Elevation")
# print(s3d)
# 
# achtal.plot <- plot_ly(x = achtal.df$results$location$lng, 
#                      y = achtal.df$results$location$lat, 
#                      z = achtal.df$results$elevation,
#                      type = "mesh3d", 
#                      #color = achtal.df$results$elevation,
#                      #colors = "YlOrRd",
#                      height = 1000)
# 
# achtal.matrix <- acast(df, longitude~latitude, value.var="elevation")
# #spread(df, latitude, elevation)
# 
# persp3D(x = df$longitude, y = df$latitude, z = achtal.matrix, main = "volcano", clab = c("height", "m"),
# breaks = seq(0,1000, by = 100))
```


