---
title: "SignBase: Maps for Swabian Jura Sites"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

# Description
This file gives code for plotting maps of the Swabian Aurignacian sites, including randomly sampled examples of sign codings for illustration purposes.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
Load the following packages. If these are not yet installed use install.packages("") to install them.
```{r}
library(ggmap)
library(gridExtra)
library(tidyverse)
library(ggpubr)
library(ggrepel)
library(googleway)
library(grid)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Load data
Run this code to load the file with data on Aurignacian objects. 
```{r}
objects <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-05-13.csv")
nrow(objects)
# Select Aurignacian objects from Swabian Jura Caves
objects <- objects[objects$archaeological_unit == "Aurignacien" &
                     objects$site_name == "Vogelherd" |
                     objects$site_name == "Hohle Fels" |
                     objects$site_name == "Geissenklösterle" |
                     objects$site_name == "Hohlenstein-Stadel", ]
nrow(objects)
```

# Overview Stats
Create data frame with object counts per site.
```{r}
sites <- count(objects, site_name)
colnames(sites) <- c("site_name", "number_of_objects")
print(sites)
```

# Collapse Factors
Collapse some factors of meta information which are uncommon and/or related. 
```{r}
# material
table(objects$material)
# object_type
table(objects$object_type)

# material: collapse manually
objects$material <- fct_collapse(objects$material,
             ivory = "ivory",
             antler = "antler",
             bone = "bone",
             other_level = "other")

# object_type: collapse manually
objects$object_type <- fct_collapse(objects$object_type,
             figurine_anthropomorph = c("figurine anthropomorph",
                          "plaquette"), 
             # note: the "plaquette" is the so-called Adorant with an anthropomorph figure on it.
             figurine_zoomorph = c("figurine zoomorph"),
             tool = c("awl", "bâton percé",
                      "rod/bâton", "spatula/lissoir"),
             tube_flute = c("flute", "tube", "possible flute", "tube/flute"),
             personal_ornament = c("personal ornament"),
             other_level = "other")
```

## Maps
Set API key. API keys need to be received for stadiamaps and the google API by signing up online.
```{r}
register_stadiamaps(key = "c56e7665-ab34-4608-8730-428e35d2766e", write = FALSE)
register_google(key = "AIzaSyDVDHFLK0KnDoxuRfEauKSzvU1Udxu92RY")
```

## Map of Germany
Map of Germany with Swabian Jura indication. 
```{r}
# Germany
germany.map.background <- get_map(location = "Germany", maptype = "stamen_toner_background", zoom = 6)
germany.map <- ggmap(germany.map.background) +
  geom_rect(aes(xmin = 9.7, xmax = 10.25, ymin = 48.35, ymax = 48.7), color = "red", fill = NA) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
germany.map
```

## Map of Swabian Jura
Plot a map of the Swabian Jura sites including Lone Valley and Ach Valley with each object and its location, and the densities of objects. Note that the object locations need to be jittered, otherwise they would just be overplotted for all objects of the same site.
```{r, fig.width = 5, fig.height = 5}
# Achtal and Lonetal, get map
map <- get_stadiamap(bbox = c(left = 9.7, bottom = 48.35, right = 10.25, top = 48.7), 
                         zoom = 12, maptype = c("stamen_terrain"), crop = TRUE, messaging = FALSE,
                         urlonly = FALSE, force = FALSE,
                         where = tempdir())
ggmap(map)

# Achtal + Lonetal Plot
object.density.map <- ggmap(map) +
  stat_density2d(data = objects,
                 aes(x = longitude, y = latitude, fill = after_stat(level)),
                 alpha = 0.1,
                 geom = "polygon") +
  geom_density2d(data = objects, 
                 aes(x = longitude, y = latitude), colour = "white") +
  scale_fill_gradient(low = "yellow", high= "red") +
  geom_point(data = objects,
             aes(jitter(longitude, amount = 0.01), jitter(latitude, amount = 0.01)), 
             size = 0.3)  +
  theme(legend.position = "none") +
  ggtitle("Swabian Jura") +
  # further segment and text annotations
  annotate("segment", x = 10.18, xend = 10, y = 48.559, yend = 48.559, colour = "black") +
  annotate("text", x = 10.05, y = 48.57, label = "Lone Valley", colour = "black") +
  annotate("segment", x = 9.77, xend = 9.85, y = 48.385, yend = 48.385, colour = "black") +
  annotate("text", x = 9.85, y = 48.395, label = "Ach Valley", colour = "black") +
  annotate("text", x = 10, y = 48.41, label = "Ulm", colour = "black") +
  annotate("segment", x = 10.18, xend = 9.77, y = 48.559, yend = 48.385, colour = "black", linetype = 3) +
  annotate("text", x = 10.02, y = 48.47, label = "c. 35 km", colour = "black")
swabianJura.map <- object.density.map +
  # add map of Germany as inset
  inset(ggplotGrob(germany.map), xmin = 9.72, xmax = 9.98, ymin = 48.5, ymax = 48.69)
swabianJura.map
```

Subset object data frame according to cave site.
```{r}
objects.gkl <- objects[objects$site_name == "Geissenklösterle", ]
objects.hfc <- objects[objects$site_name == "Hohle Fels", ]
objects.vhc <- objects[objects$site_name == "Vogelherd", ]
objects.hss <- objects[objects$site_name == "Hohlenstein-Stadel", ]
```

## Maps of Achtal and Lonetal

### Downsample the object data
This is necessary since all the objects would overcrowed the plots.
```{r}
set.seed(15022018)
# Geissenklösterle
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.gkl.sample <- objects.gkl[objects.gkl$object_type != "other", ]
# sample subset of objects randomly (otherwise the plot is overcrowded)
objects.gkl.sample <- sample_n(objects.gkl.sample, 20)

# Hohle Fels
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.hfc.sample <- objects.hfc[objects.hfc$object_type != "other", ]
# sample subset of objects randomly (otherwise the plot is overcrowded)
objects.hfc.sample <- sample_n(objects.hfc.sample, 20)

# Vogelherd 
# Remove objects of type "other", this reduces the number of objects to display strings for
objects.vhc.sample <- objects.vhc[objects.vhc$object_type != "other", ]
# sample subset objects randomly (otherwise the plot is overcrowded)
objects.vhc.sample <- sample_n(objects.vhc.sample, 20)
```

### Maps with objects for Achtal and Lonetal

Plot a map of the Lonetal.
```{r, warning = F}
# Lonetal, get map
map.lonetal <- get_map(location = c(lon = 10.185, lat = 48.5546), 
                         zoom = 15, maptype = "satellite", source = "google")
ggmap(map.lonetal)

# Lonetal plot
object.map.lonetal <- ggmap(map.lonetal) +
  # Hohlenstein-Stadel
  geom_text_repel(data = objects.hss, 
                  aes(x = longitude, y = latitude, 
                      label = paste(object_id, substring(sign_coding, 1, 30), sep = "  ")),
                  color = "white",
                  direction = 'y',
                  fontface = "bold",
                  nudge_x = 0.004,
                  hjust = 0,
                  alpha = 0.8,
                  segment.size = 0.4,
                  size = 3,
                  max.overlaps = Inf) +
  geom_segment(aes(x = 10.194, y = 48.559, xend = 10.17255, yend = 48.54975), color = "white",
               linetype = 3) +
  theme(legend.position = "none") +
  ggtitle("Lone Valley") +
  annotate("text", x = 10.185, y = 48.554, label = "1.9 km", colour = "white") +
  annotate("segment", x = 10.194, xend = 10.188, y = 48.559, yend = 48.559, colour = "white") +
  annotate("text", x = 10.188, y = 48.56, label = "Vogelherd", colour = "white") +
  annotate("segment", x = 10.17255, xend = 10.17255, y = 48.54975, yend = 48.547, colour = "white") +
  annotate("text", x = 10.178, y = 48.547, label = "Hohlenstein-Stadel", colour = "white")
object.map.lonetal
```

Map of the Achtal
```{r, warning = F}
# Achtal, get map
map.achtal <- get_map(location = c(lon = 9.76473, lat = 48.39075), 
                         zoom = 14, maptype = "satellite", source = "google")
ggmap(map.achtal)

# Achtal Plot
object.map.achtal <- ggmap(map.achtal) +
  ggtitle("Ach Valley") +
  annotate("segment", x = 9.772222, y = 48.398333, xend = 9.754103, yend = 48.379117, color = "white",
               linetype = 3) +
  annotate("text", x = 9.768, y = 48.389, label = "2.5 km", colour = "white") +
  annotate("segment", x = 9.772222, xend = 9.772222, y = 48.398333, yend = 48.403, colour = "white") +
  annotate("text", x = 9.772222, y = 48.404, label = "Geissenklösterle", colour = "white") +
  annotate("segment", x = 9.754103, xend = 9.754103, y = 48.379117, yend = 48.376, colour = "white") +
  annotate("text", x = 9.754103, y = 48.375, label = "Hohle Fels", colour = "white")
object.map.achtal
```

Map with objects for Vogelherd cave.
```{r, warning = F}
# Vogelherd, get map
map.vogelherd <- get_map(location = c(lon = 10.195, lat = 48.5586), 
                         zoom = 19, maptype = "satellite", source = "google")
ggmap(map.vogelherd)

# Vogelherd plot
object.map.vogelherd <- ggmap(map.vogelherd) +
  geom_text_repel(data = objects.vhc.sample, 
                  aes(x = 10.1943, y = 48.5586, 
                      label = paste(object_id, substring(sign_coding, 1, 30), sep = "  ")),
                  color = "white",
                  fontface = "bold",
                  alpha = 0.8,
                  direction = 'y',
                  nudge_x = 0.0003,
                  hjust = 0,
                  #segment.size = 0.4,
                  size = 3,
                  max.overlaps = Inf) +
  ggtitle(paste("Vogelherd Cave, total n = ", nrow(objects.vhc), sep ="")) +
  theme(legend.position = "bottom")
object.map.vogelherd
```

Map with objects for Geissenklösterle cave.
```{r, warning = F}
# Geissenklösterle, get map
map.gkl <- get_map(location = c(lon = 9.7725, lat = 48.3982), 
                         zoom = 18, maptype = "satellite", source = "google")
ggmap(map.gkl)

# Geissenklösterle plot
object.map.gkl <- ggmap(map.gkl) +
  geom_text_repel(data = objects.gkl.sample, 
                  aes(x = 9.7715, y = 48.3982, 
                      label = paste(object_id, substring(sign_coding, 1, 30), sep = "  ")),
                  direction = 'y',
                  color = "white",
                  fontface = "bold",
                  nudge_x = 0.0004,
                  hjust = 0,
                  #segment.size = 0.4,
                  size = 3,
                  max.overlaps = Inf) +
  ggtitle(paste("Geissenklösterle Cave, total n = ", nrow(objects.gkl), sep ="")) +
  theme(legend.position = "none")
object.map.gkl
```
Map with objects for Hohle Fels cave.
```{r, warning = F}
# Hohle Fels, get map
map.hfc <- get_map(location = c(lon = 9.7547, lat = 48.379117), 
                         zoom = 19, maptype = "satellite", source = "google")
ggmap(map.hfc)

# Hohle Fels plot
object.map.hfc <- ggmap(map.hfc) +
  geom_text_repel(data = objects.hfc.sample, 
                  aes(x = 9.754103, y = 48.379117, color = object_type, 
                      label = paste(object_id, substring(sign_coding, 1, 30), sep = "  ")),
                  direction = 'y',
                  color = "white",
                  fontface = "bold",
                  nudge_x = 0.0004,
                  hjust = 0,
                  #segment.size = 0.4,
                  size = 3,
                  max.overlaps = Inf) +
  ggtitle(paste("Hohle Fels Cave, total n = ", nrow(objects.hfc), sep ="")) +
  theme(legend.position = "none")
object.map.hfc
```

Combine plots.
```{r, fig.width = 15, fig.height = 5, warning = F}
combined.plot <- ggarrange(swabianJura.map, object.map.lonetal, object.map.vogelherd,
                           object.map.achtal, object.map.hfc, object.map.gkl,
                           ncol = 3, nrow = 2,
                           labels = "AUTO")
combined.plot
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/maps/maps_combined.pdf", 
       combined.plot, dpi = 500, width = 15, height = 10, device = cairo_pdf)
```

# Package References
```{r, echo = F}
print(citation("ggmap"), style = "text")
```

```{r, echo = F}
print(citation("gridExtra"), style = "text")
print(citation("tidyverse"), style = "text")
print(citation("ggpubr"), style = "text")
print(citation("ggrepel"), style = "text")
print(citation("googleway"), style = "text")
print(citation("grid"), style = "text")
```
