---
title: "SignBase: Maps for Aurignacian Sites"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
R version: R version 3.4.4 (2018-03-15) -- "Someone to Lean On"
Linux version: Ubuntu 18.04.3 LTS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

# Load libraries
Load the following packages. If these are not yet installed use install.packages("") to install them.
```{r}
library(ggmap)
library(plyr)
library(gridExtra)
```

# Load data
Run this code to load the file with data on Aurignacian objects.
```{r}
objects <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-04-19.csv")
nrow(objects)
# Select Aurignacian objects from Swabian Jura Caves
objects <- objects[objects$archaeological_unit == "Aurignacien" &
                     objects$site_name == "Vogelherd" |
                     objects$site_name == "Hohle Fels" |
                     objects$site_name == "Geissenklösterle" |
                     objects$site_name == "Hohlenstein-Stadel", ]
nrow(objects)
```
# Preprocessing

Create data frame with object counts per site.
```{r}
sites <- count(objects$site_name)
colnames(sites) <- c("site_name", "number_of_objects")
```

Merge sites and objects.short
```{r}
sites <- merge(sites, objects.short, by = "site_name")
```

## ggmap with OBJECT locations
Set API key. 
```{r}
register_stadiamaps(key = "c56e7665-ab34-4608-8730-428e35d2766e", write = FALSE)
```

Load stamenmap of Europe. 
```{r}
map <- get_stadiamap(bbox = c(left = 9.7, bottom = 48.35, right = 10.25, top = 48.6), 
                         zoom = 12, maptype = c("stamen_terrain"), crop = TRUE, messaging = FALSE,
                         urlonly = FALSE, force = FALSE,
                         where = tempdir())
ggmap(map)
```

Plot a map of Europe with each object and its location, and the densities of objects. Note that the object locations need to be jittered, otherwise they would just be overplotted for all objects of the same site.
```{r}
object.density.map <- ggmap(map) +
  stat_density2d(data = objects,
                 aes(x = longitude, y = latitude, fill = ..level..),
                 alpha = 0.3, size = 0.1,
                 geom = "polygon") +
  geom_density2d(data = objects, 
                 aes(x = longitude, y = latitude), size = 0.3, colour = "white") +
  scale_fill_gradient(low = "yellow", high= "red") +
  geom_point(data = objects,
             aes(jitter(longitude, amount = 0.005), jitter(latitude, amount = 0.005)), 
             size = 0.2)  +
  theme(legend.position = "none") 
  #geom_text(x = -3, y = 51.5, label = "b)", size = 7) +
  #annotate("segment", x = 10, xend = 12.2, y = 48.5, yend = 46, colour = "black") +
  #annotate("text", x = 16, y = 46, label = "Swabian Jura", colour = "black")
object.density.map
```

## Put both maps together in one panel plot
```{r, fig.width = 8, fig.height = 3}
density.maps <- grid.arrange(site.density.euromap, object.density.euromap, ncol = 2)
```

# Safe as pdf
Optionally, this plot can be safed to a file as pdf via:
```{r, fig.width = 8, fig.height = 3}
ggsave("Figures/Figure_densityMaps.pdf", density.maps, dpi = 300, scale = 1, device = cairo_pdf)
```

