---
title: "Analyses Randomization"
author: "Chris Bentz"
date: "2/27/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggExtra)
library(ggpubr)
library(gridExtra)
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object. 
```{r}
# load signbase estimations for strings on objects
sign.data <- read.csv("~/Github/UpperPalCombinatorics/results/signBase_features.csv")
# select Aurignacian objects
sign.data <- sign.data[sign.data$archaeological_unit == "Aurignacien", ] 
```

# Plot

## Entropy rate vs. Unigram entropy 
Entropy rate vs. unigram entropy for the original versus randomized sequences.
```{r, fig.width = 6, fig.height = 6, warning = FALSE}
hrate.random.plot <- ggplot(sign.data, aes(x = hrate.chars, y =  hrate.chars.random, 
                                    size = num.char)) + 
  geom_point(alpha = 0.3) +
  theme(legend.position = c(.8, .25)) +
  labs(x = " Entropy rate for characters (original)", y = "Entropy rate for characters (randomized)")
hrate.random.plot <- ggMarginal(hrate.random.plot,
                                  type = "density")
hrate.random.plot
```

## Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_signBase_random.pdf", 
       hrate.random.plot, dpi = 300, width = 6, height = 6, device = cairo_pdf)
```

# Correlation
```{r}
cor(sign.data$hrate.chars, sign.data$hrate.chars.random, method = "pearson")
```
