---
title: "Plots with Feature Values"
author: ""
date: "22/02/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(ggExtra)
library(gridExtra)
library(ggpubr)
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object. Comparative sample: A comparative sample of feature values for languages across the world, as well as other symbolic and non-symbolic sign sequences, is loaded here too. Note that the feature values were calculated as part of another Github project on ``Natural Language Fingerprints'' (NaLaFi), and published in Bentz (2023). 
```{r}
# SIGNBASE
# load signbase estimations for strings on objects
sign.data <- read.csv("~/Github/UpperPalCombinatorics/results/signBase_features.csv")

# select columns for direct comparison with NaLaFi sample
sign.data <- select(sign.data, c('archaeological_unit', 'num.char', 'huni.chars', 'hrate.chars', 
                                 'ttr.chars', 'rm.chars'))
# change column names
colnames(sign.data) <- c('subcorpus','num.char','huni.chars', 'hrate.chars', 
                         'ttr.chars', 'rm.chars')

# CDLI
# read feature values for proto-cuneiform from CDLI
# Uruk V
procuneiV.data <- read.csv("~/Github/UpperPalCombinatorics/results/procunei_urukV_features.csv")

# add column with subcorpus information
procuneiV.data$subcorpus <- rep("Proto-Cuneiform (Uruk V)", nrow(procuneiV.data))

# select columns for direct comparison with NaLaFi sample
procuneiV.data <- select(procuneiV.data, c('subcorpus', 'num.char', 'huni.chars', 'hrate.chars', 
                                 'ttr.chars', 'rm.chars'))
# Uruk IV
procuneiIV.data <- read.csv("~/Github/UpperPalCombinatorics/results/procunei_urukIV_features.csv")

# add column with subcorpus information
procuneiIV.data$subcorpus <- rep("Proto-Cuneiform (Uruk IV)", nrow(procuneiIV.data))

# select columns for direct comparison with NaLaFi sample
procuneiIV.data <- select(procuneiIV.data, c('subcorpus', 'num.char', 'huni.chars', 'hrate.chars', 
                                 'ttr.chars', 'rm.chars'))

# NALAFI
# read feature values for languages and other sign systems
nalafi <- read.csv("~/Github/NaLaFi/results/features.csv")
# select columns
nalafi <- select(nalafi, c('subcorpus','num.char','huni.chars', 'hrate.chars', 
                           'ttr.chars', 'rm.chars'))
```

# Preprocessing
```{r}
# select only data with 100 chars
nalafi <- nalafi[nalafi$num.char == 100, ]

# combine the data frames
estimations.df <- rbind(sign.data, procuneiV.data, procuneiIV.data, nalafi)
nrow(estimations.df)

# select sign strings of certain length
# estimations.df <- estimations.df[estimations.df$num.char > 10,]

# select only certain subcorpora for plotting
estimations.df.selected <- estimations.df[estimations.df$subcorpus == "Aurignacien"|
                                          #estimations.df$subcorpus == "Aurignacien/Gravettian"|
                                          estimations.df$subcorpus == "Proto-Cuneiform (Uruk V)"|
                                          estimations.df$subcorpus == "Proto-Cuneiform (Uruk IV)"|  
                                          estimations.df$subcorpus == "udhr", ]
```

## Entropy rate vs. unigram entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y = hrate.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  #geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  theme(legend.position = c(.8,.2))
  #facet_wrap(~subcorpus)
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
huni.hrate.chars.plot
```

## TTR vs. repetition rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  #geom_density2d(alpha = .5) +
  labs(x = "TTR for characters", y = "Repetition rate for characters") +
  theme(legend.position = "none")
  #facet_wrap(~subcorpus)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
ttr.rm.chars.plot
```

Combine figures.
```{r}
combined.plot <- grid.arrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2)
```

## Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_chars.pdf", combined.plot, dpi = 300, width = 20, height = 10, device = cairo_pdf)
```


