---
title: "Plots with Feature Values"
author: ""
date: "15/11/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(ggExtra)
library(ggpubr)
```

# Load Data
Load csv file with quantitative feature values for samples per row. Comparative sample: A comparative sample of feature values for languages across the world, as well as other symbolic and non-symbolic sign sequences, is loaded here too. Note that the feature values were calculated as part of another Github project on ``Natural Language Fingerprints'' (NaLaFi), and published in Bentz (2023). 
```{r}
upperPal <- read.csv("~/Github/UpperPalCombinatorics/results/features.csv")
# select columns
upperPal <- select(upperPal, c('period','num.char','huni.chars', 'hrate.chars', 'ttr.chars', 'rm.chars'))
# change column names
colnames(upperPal) <- c('subcorpus','num.char','huni.chars', 'hrate.chars', 'ttr.chars', 'rm.chars')

# read feature values for languages and other sign systems
nalafi <- read.csv("~/Github/NaLaFi/results/features.csv")
# select columns
nalafi <- select(nalafi, c('subcorpus','num.char','huni.chars', 'hrate.chars', 'ttr.chars', 'rm.chars'))
# select only data with 100 chars
nalafi <- nalafi[nalafi$num.char == 100, ]

# combine the data frames
estimations.df <- rbind(upperPal, nalafi)
nrow(estimations.df)

# select only certain subcorpora for plotting
estimations.df.selected <- estimations.df[estimations.df$subcorpus == "aurignacian"|
                                          estimations.df$subcorpus == "magdalenian"|
                                          estimations.df$subcorpus == "teddi", ]
```
## Entropy rate vs. unigram entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y = hrate.chars, 
                                    colour = subcorpus, fill = subcorpus)) + 
  geom_point(alpha = 0.3, size  = 1) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") #+
  #theme(legend.position = c(.8,.2))
  #facet_wrap(~subcorpus)
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, type = "histogram", colour = "white", 
                                      fill = "light grey")
huni.hrate.chars.plot
```


## TTR vs. repetition rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = subcorpus)) + 
  geom_point(alpha = 0.3, size  = 1) +
  labs(x = "TTR for characters", y = "Repetition rate for characters") #+
  #theme(legend.position = c(.8,.2))
  #facet_wrap(~subcorpus)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, type = "histogram", colour = "white", 
                                      fill = "light grey")
ttr.rm.chars.plot
```

## Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/huni_hrate_chars.pdf", huni.hrate.chars.plot, dpi = 300, width = 5, height = 5, device = cairo_pdf)
```


