---
title: "Preprocessing"
author: ''
date: "11/1/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables preprocessing of paleolithic sign sequences from the Swabian Aurignacian and the Magdalenian ``necklace'' of the Saint-Germain-la-Rivière burial.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message=F}
library(dplyr)
```

# Preprocessing

## Magdalenian
Preprocessing and storage of cleaned file for signs of the Magdalenian ``necklace''.
Load data from github repo and remove missing rows.
```{r}
# load data from github repo
magda.data <- read.csv(
  "~/Github/UpperPalCombinatorics/data/magdalenian/StGermain_original.csv")
# select columns
magda.data <- select(magda.data, c('VanhaerenDErrico2003_Fig.6', 'DutkiewiczBentz2023',
                                   'lateralization','position', 'coding', 'certainty'))
# remove rows for which there is no sign coding
magda.data <- magda.data[!magda.data$coding == "", ]
```

Remove UTF8 characters which do not represent actual signs on the objects, i.e "punctuation".
```{r}
# remove some special characters 
coding.clean <- gsub('[!?]', '', magda.data$coding)
# remove sign strings which are not discrete, i.e. overlapping sign types 
# -> indicated by parentheses '()' 
coding.clean <- gsub("\\([^()]*\\)", "", coding.clean)
# remove sign strings which are not in linear order, i.e. those grouped in visual fields 
# -> indicated by square brackets '[]' 
coding.clean <- gsub("\\[(.*?)\\]", "", coding.clean)
# print for visual inspection
# print(coding.clean)

# add column to original data frame again
magda.data$coding_clean <- coding.clean
```

Safe to file.
```{r}
write.csv(magda.data, 
          file = "~/Github/UpperPalCombinatorics/data/magdalenian/StGermain_clean.csv",
          row.names = F)
```

## Aurignacian 
Preprocessing and storage of cleaned file for the signs of the mobile objects from the Swabian Jura.
Load data from github repo and remove missing rows.
```{r}
# load data from github repo
aur.data <- read.csv(
  "~/Github/UpperPalCombinatorics/data/aurignacian/signBase_Version1.0_CodingsUTF8.csv")
# select columns
aur.data <- select(aur.data, c('object_id', 'site_name', 'date_bp_max_min',
                                   'material','object_type', 'coding'))
# remove rows for which there is no sign coding
aur.data <- aur.data[!aur.data$coding == "", ]
```

Remove UTF8 characters which do not represent actual signs on the objects, i.e "punctuation".
```{r}
# remove special characters
coding.clean <- gsub('[!?]', '', aur.data$coding)
# remove sign strings which are not discrete, i.e. overlapping sign types 
# -> indicated by parentheses '()' 
coding.clean <- gsub("\\([^()]*\\)", "", coding.clean)
# remove sign strings which are not in linear order, i.e. those grouped in visual fields 
# -> indicated by square brackets '[]' 
coding.clean <- gsub("\\[(.*?)\\]", "", coding.clean)
# print for visual inspection
#print(coding.clean)

# add column to original data frame again
aur.data$coding_clean <- coding.clean
```

Safe to file.
```{r}
write.csv(aur.data, 
          file = "~/Github/UpperPalCombinatorics/data/aurignacian/swabianJura_clean.csv",
          row.names = F)
```
