---
title: "Preprocessing"
author: ''
date: "16/2/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables preprocessing of upper-paleolithic sign sequences as published at https://www.signbase.org. The csv file with latest objects and sign codings can be downloaded from there.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message=F}
#library(dplyr)
```

# Preprocessing
Preprocessing and storage of cleaned file for signs. Load data from github repo and remove missing rows.
```{r}
# load data from github repo
sign.data <- read.csv(
  "~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-02-16.csv")
# remove certain columns
sign.data <- subset(sign.data, select = -c(id, comments, short_description, references))
# remove rows for which there is no sign coding
sign.data <- sign.data[!sign.data$sign_coding == "", ]
```

Remove UTF8 characters which do not represent actual signs on the objects, i.e "punctuation".
```{r}
# remove some special characters 
coding.clean <- gsub('[!?]', '', sign.data$sign_coding)
# remove sign strings which are not discrete, i.e. overlapping sign types 
# -> indicated by parentheses '()' 
coding.clean <- gsub("\\([^()]*\\)", "", coding.clean)
# remove sign strings which are not in linear order, i.e. those grouped in visual fields 
# -> indicated by square brackets '[]' 
coding.clean <- gsub("\\[(.*?)\\]", "", coding.clean)
# replace multiple underscores by single underscore
coding.clean <- gsub("( _ )+", " _ ", coding.clean)

# add column to original data frame again
sign.data$coding_clean <- coding.clean
```

Safe to file.
```{r}
write.csv(sign.data, 
          file = "~/Github/UpperPalCombinatorics/data/signBase/signBase_clean.csv",
          row.names = F)
```

