---
title: "Randomization"
author: ''
date: "15/1/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables randomization of sign strings which occur on different parts of an object (indicated by underscore `_'). We do not know the order of these strings (if any), hence they are randomized. In the case of the Magdalenian teeth of the Saint-Germain-la-Rivière burial, we consider them as one assemblage, that is, we concatenate them by also using underscores.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message=F}
#library(stringr)
```

# Load Data
```{r}
mag <- read.csv("~/Github/UpperPalCombinatorics/data/magdalenian/StGermain_clean.csv")
head(mag)
aur <- read.csv("~/Github/UpperPalCombinatorics/data/aurignacian/swabianJura_clean.csv")
head(aur)
```

# Randomize and Concatenate

## Magdalenian burial teeth
Concatenate and randomize the sign strings on Magdalenian teeth.
```{r}
# choose observations based on "certainty" factor 
mag <- mag[mag$certainty == "very certain" |
            mag$certainty == "certain" ,]
# add length in characters
mag$nchar <- nchar(mag$coding_clean)
# get average number of signs on teeth
mean(mag$nchar) # ca. 4

# choose number of random samples
n = 100
# initialize data frame
mag.sample <- data.frame(id = numeric(0), coding_sample = character(0))
# run loop to randomize, concatenate and safe in data frame
for (i in 1:n){
  # random sampling, choose teeth randomly, the size is here chosen
  # to give c. 100 signs (avg. 4 per tooth) per random sample 
  coding_sample <- sample(mag$coding_clean, size = 25)
  # concatenations using underscore
  coding_sample <- paste(coding_sample, collapse = "_")
  id <- i
  # store in local data frame
  local.df <- data.frame(id, coding_sample)
  # append to global data frame
  mag.sample <- rbind(mag.sample, local.df)
}
# safe to file
write.csv(mag.sample, "~/Github/UpperPalCombinatorics/data/magdalenian/StGermain_randomized.csv", row.names = F)
```

## Simple Stats
Simple statistics for the concatenated sign strings on Magdalenian teeth.
```{r}
# number of characters
nchar(mag.sample$coding_sample[1])
```

## Swabian Jura
Randomize and concatenate sign strings on objects of the Swabian Jura.
```{r}
# add length in characters
aur$nchar <- nchar(aur$coding_clean)
# exclude empty rows for codings
aur <- aur[aur$nchar > 0, ]
# get average number of signs on objects
mean(aur$nchar) # ca. 14

# randomize and concatenate sign strings
# choose number of randomizations
n = 100
# initialize data frame
aur.sample <- data.frame(id = numeric(0), coding_sample = character(0))
# run loop to randomize, concatenate and safe in data frame
for (i in 1:n){
  # random sampling
  # note that 7 strings are here chosen since this will give
  # an expected length of c. 100 signs
  coding_sample <- sample(aur$coding_clean, size = 7)
  # concatenations using underscore
  coding_sample <- paste(coding_sample, collapse = "_")
  # replace multiple underscores by single underscore
  coding_sample <- gsub("( _ )+", "_", coding_sample)
  # get id
  id <- i
  # store in local data frame
  local.df <- data.frame(id, coding_sample)
  # append to global data frame
  aur.sample <- rbind(aur.sample, local.df)
}
# safe to file
write.csv(aur.sample, "~/Github/UpperPalCombinatorics/data/aurignacian/swabianJura_randomized.csv", row.names = F)
```


