---
title: "Randomization"
author: ''
date: "16/02/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables concatenation and randomization of sign strings which occur on different parts of an object (indicated by underscore `_'). We do not know the order of these strings (if any), hence, if we want to combine them, they need to be randomized.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message = F}
#library(stringr)
```

# Load Data
```{r}
sign.data <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/signBase_clean.csv")
head(sign.data)
```

# Randomize
Randomize sign sequences per object.
```{r}
# split sequences on '_'


# randomize and concatenate sign strings
# choose number of randomizations
n = 100
# initialize data frame
aur.sample <- data.frame(id = numeric(0), coding_sample = character(0))
# run loop to randomize, concatenate and safe in data frame
for (i in 1:n){
  # random sampling
  # note that 7 strings are here chosen since this will give
  # an expected length of c. 100 signs
  coding_sample <- sample(aur$coding_clean, size = 7)
  # concatenations using underscore
  coding_sample <- paste(coding_sample, collapse = "_")
  # replace multiple underscores by single underscore
  coding_sample <- gsub("( _ )+", "_", coding_sample)
  # get id
  id <- i
  # store in local data frame
  local.df <- data.frame(id, coding_sample)
  # append to global data frame
  aur.sample <- rbind(aur.sample, local.df)
}
# safe to file
write.csv(aur.sample, "~/Github/UpperPalCombinatorics/data/aurignacian/swabianJura_randomized.csv", row.names = F)
```


