---
title: "Randomization Analyses"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file provides code to estimate to what extend randomized sign sequences deviate from the original sequences given in SignBase in terms of the entropy rates. This is necessary since the order of strings into sequences (strings are delimited by underscore in SignBase) is arbitrary. Note that randomization does only affect the entropy rate calculated over the sign tokens in sequences. All the other measures (TTR, unigram entropy, repetition measure) yield the same values for randomized sequences.  

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggExtra)
library(ggpubr)
library(gridExtra)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object. 
```{r}
# load signbase estimations for strings on objects
sign.data <- read.csv("~/Github/PaleoSigns/results/features/signBase_features.csv")
# select Aurignacian objects
sign.data <- sign.data[sign.data$archaeological_unit == "Aurignacien", ] 
```

# Correlation
```{r}
cor.value <- round(cor(sign.data$hrate.chars, sign.data$hrate.chars.random, method = "pearson"), digits = 2)
cor.pvalue <- round(cor.test(sign.data$hrate.chars, sign.data$hrate.chars.random, method = "pearson")$p.value, digits = 2)
sample.size <- nrow(sign.data)
print(cor.value)
print(cor.pvalue)
print(sample.size)
```

# Plot

## Entropy rate 
Entropy rate for the original versus randomized sequences (i.e. strings of sign tokens where randomized).
```{r, fig.width = 6, fig.height = 6, warning = FALSE}
hrate.random.plot <- ggplot(sign.data, aes(x = hrate.chars, y =  hrate.chars.random)) + 
  geom_point(aes(size = num.tokens), alpha = 0.3) +
  geom_smooth(method = "loess") +
  theme(legend.position = c(.8, .25)) +
  annotate("text", x = c(0.5,0.5,0.5), y = c(1.7,1.6,1.5), 
           label = c(paste("r = ", cor.value, sep =""), 
                     "p-value < 0.0001",
                     paste("n = ", sample.size, sep ="")), size = 4) +
  labs(x = " Entropy rate for tokens (original)", 
       y = "Entropy rate for tokens (randomized)") +
  guides(size = guide_legend(title="No. of signs"))
hrate.random.plot <- ggMarginal(hrate.random.plot,
                                  type = "density")
hrate.random.plot
```

## Safe complete figure to file
```{r}
ggsave("~/Github/PaleoSigns/figures/combinedPlot_signBase_random.pdf", 
       hrate.random.plot, dpi = 300, width = 6, height = 6, device = cairo_pdf)
```

# Package References
```{r, echo = F}
print(citation("ggplot2"), style = "text")
print(citation("ggExtra"), style = "text")
print(citation("ggpubr"), style = "text")
print(citation("gridExtra"), style = "text")
```
