---
title: "Chinese Characters vs. Strokes"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file provides functionality for plotting feature value distributions in density plots. This file in particular focuses on the difference between Chinese complex characters and individual strokes. See the file preprocess_TeDDi.Rmd for further details on the mapping.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggrepel)
library(ggExtra)
library(ggpubr)
library(ggridges)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Load Data
Load csv file with quantitative feature values for a comparative sample of feature values for languages across the world. This is derived from the Text Data Diversity (TeDDi) sample (Moran et al 2022). 

```{r}
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features/teddi_features.csv")
head(estimations.df)
```

# Preprocessing
```{r}
# select only texts written in Han simplified (Hans)
estimations.df.selected <- estimations.df[estimations.df$writing_system == "Hans" |
                                       estimations.df$writing_system == "Hans (strokes)", ]
```

# Density Plots

## Length
Density plot for lengths in number of characters.
```{r}
length.plot <- ggplot(estimations.df.selected, aes(x = num.char, y = writing_system, 
                                    fill = writing_system)) + 
  # add density ridges with quantiles = 2, i.e. median as a line
  geom_density_ridges(alpha = 0.5, quantile_lines = TRUE, quantiles = 2) +
  # limit the x-range 
  xlim(0, 100) +
  xlab("Number of Characters/Strokes") +
  theme(legend.position = "none")
length.plot
```

Get subset of data points to plot labels.
```{r}
# how many data points with a given sequence length?
table(estimations.df.selected$num.char)

# choose subset of data points for labeling
labels.df <- estimations.df.selected[estimations.df.selected$num.char > 10 &
                                       estimations.df.selected$num.char < 20, ]
```

## Entropy rate vs. Unigram Entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y =  hrate.chars, 
                                                             color = writing_system, size = num.char)) + 
  geom_point(alpha = 0.3) +
  xlim(-0.2, 5) +
  # geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters/strokes", y = "Entropy rate for characters/strokes") +
  geom_label_repel(data = labels.df,
                  aes(label = text),
                  size = 3,
                  hjust = 0,
                  direction = 'y',
                  fontface = "bold",
                  nudge_x = 1,
                  max.overlaps = Inf,
                  show.legend = FALSE) +
    theme(legend.position = c(.2, .75)) +
  guides(size = guide_legend(title = "No. of chars/strokes"))
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#huni.hrate.chars.plot
```

## TTR vs. Repetition Rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = writing_system, size = num.char)) + 
  geom_point(alpha = 0.3) +
  xlim(0, 1.5) +
  ylim(-0.05, 1) +
  theme(legend.position = "none") +
  labs(x = "TTR for characters/strokes", y = "Repetition rate for characters/strokes") +
  geom_label_repel(data = labels.df,
                   aes(label = text),
                   size = 3,
                   hjust = 0,
                   direction = 'y',
                   fontface = "bold",
                   nudge_x = 0.2,
                   max.overlaps = Inf,
                   show.legend = FALSE)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#ttr.rm.chars.plot
```

Combine figures.
```{r, fig.width = 15, fig.height = 7.5}
combined.plot <- ggarrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2,
                               labels = c("A", "B"))
combined.plot
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_chineseStrokes.pdf", 
       combined.plot, dpi = 300, width = 15, height = 7.5, device = cairo_pdf)
```

# Package References
```{r, echo = F}
print(citation("ggrepel"), style = "text")
print(citation("ggExtra"), style = "text")
print(citation("ggpubr"), style = "text")
print(citation("ggridges"), style = "text")
```

# Literature References 

Jolliffe, I. T., & Cadima, J. (2016). Principal component analysis: a review and recent developments. Philosophical transactions of the royal society A: Mathematical, Physical and Engineering Sciences, 374(2065), 20150202.

Moran, S., Bentz, C., Gutierrez-Vasques, X., Pelloni, O., & Samardzic, T. (2022, June). TeDDi sample: text data diversity sample for language comparison and multilingual NLP. In Proceedings of the Thirteenth Language Resources and Evaluation Conference (pp. 1150-1158).
