---
title: "Analyses: Feature Value Distributions"
author: ""
date: "28/03/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file provides functionality for plotting feature value distributions in density plots. It also contains code for PCA analyses given the values of the four statistical features (entropy rate, unigram entropy, TTR, and repetition rate).

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
#library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)
library(stringr)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(ggExtra)
library(gridExtra)
library(ggpubr)
library(ggridges)
library(GGally)
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object (paleolithic signs) and tablet (cuneiform). Comparative sample: A comparative sample of feature values for languages across the world is derived from the Text Data Diversity (TeDDi) sample (Moran et al 2022). 

```{r}
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features/features_combined.csv")
head(estimations.df)
```

# Preprocessing
```{r}
# select sign strings of certain length
#estimations.df <- estimations.df[estimations.df$num.char > 1,]

# select only certain subcorpora for plotting
estimations.df.selected <- estimations.df[estimations.df$subcorpus == "Aurignacien"|
                                          #estimations.df$subcorpus == "Aurignacien/Gravettian"|
                                          estimations.df$subcorpus == "Cuneiform (Uruk V)"|
                                          estimations.df$subcorpus == "Cuneiform (Uruk IV)"| 
                                          estimations.df$subcorpus == "Cuneiform (Uruk III)"|  
                                          estimations.df$subcorpus == "TeDDi", ]
```

# Simple Stats

```{r}
# number of sequences
nrow(estimations.df.selected)
# number of rows (i.e. sequences) per subcorpus
table(estimations.df.selected$subcorpus)
# get mean values of lengths in characters per subcorpus
aggregate(estimations.df.selected$num.char~
            estimations.df.selected$subcorpus, FUN = mean)
# get median values of lengths in characters per subcorpus
aggregate(estimations.df.selected$num.char~
            estimations.df.selected$subcorpus, FUN = median)
```

# Density Plots

## Length
Density plot for lengths in number of characters.
```{r, fig.width = 2, fig.height = 5}
length.plot <- ggplot(estimations.df.selected, aes(x = num.char, y = subcorpus, 
                                    fill = subcorpus)) + 
  # add density ridges with quantiles = 2, i.e. median as a line
  geom_density_ridges(alpha = 0.5, quantile_lines = TRUE, quantiles = 2) +
  # limit the x-range 
  xlim(0, 100) +
  xlab("Number of Characters") +
  theme(legend.position = "none")
#length.plot
```

Safe complete figure to file
```{r}
#ggsave("~/Github/UpperPalCombinatorics/figures/lengthPlot_chars.pdf", 
#       length.plot, dpi = 300, width = 5, height = 5, device = cairo_pdf)
```

Get subset of data points to plot labels.
```{r}
# how many data points with a given sequence length?
table(estimations.df.selected$num.char)

# choose subset of data points for labeling
labels.df <- estimations.df.selected[estimations.df.selected$num.char > 9 &
                                       estimations.df.selected$num.char < 11 &
                                       #estimations.df.selected$ttr.chars < 1 &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk V)" &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk IV)" &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk III)", ]
```

## Entropy rate vs. Unigram Entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y =  hrate.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  xlim(0, 5) +
  #ylim(-0.5, 4.5) +
  # geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  geom_label_repel(data = labels.df,
                  aes(label = sequence),
                  size = 3,
                  max.overlaps = Inf,
                  show.legend = FALSE) +
    theme(legend.position = c(.2, .75))
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#huni.hrate.chars.plot
```

## TTR vs. Repetition Rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  theme(legend.position = "none") +
  labs(x = "TTR for characters", y = "Repetition rate for characters")
  # geom_label_repel(data = labels.df, 
  #                  aes(label = sequence),
  #                  position = position_nudge_repel(x = 0.25, y = 0),
  #                  size = 3,
  #                  max.overlaps = Inf,
  #                  show.legend = FALSE)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#ttr.rm.chars.plot
```

Combine figures.
```{r, fig.width = 20, fig.height = 10}
combined.plot1 <- ggarrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2,
                               labels = c("A", "B"))
```

Safe complete figure to file
```{r}
#ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_chars.pdf", 
#       combined.plot1, dpi = 300, width = 20, height = 10, device = cairo_pdf)
```

# Correlation Plot
Correlation plots for the statistical feature values.
```{r, fig.width = 12, fig.height = 12}
cor.plot <- ggpairs(estimations.df.selected, columns = c("num.char", "huni.chars", "hrate.chars",
                               "ttr.chars","rm.chars"), 
        aes(color = subcorpus,
            alpha = 0.5))
cor.plot
```

Safe complete figure to file.
```{r}
#ggsave("~/Github/UpperPalCombinatorics/figures/correlation_plot.pdf", 
#       cor.plot, dpi = 300, width = 12, height = 12, device = cairo_pdf)
```


# Principal Component Analysis (PCA)
Since we have overall four dimensions of statistical values derived from the sign sequences (unigram entropy, entropy rate, TTR, repetition rate), we can first use a principal component analysis to reduce the number of dimensions to the first two principal components. This allows us to see the clustering of data points in two dimensional space. 

Note that we here use PCA as a data visualization and exploration tool, rather than in an inferential sense (cf. Jolliffe and Cadima, 2016, p. 2). For drawing statistical inferences we rather employ statistical analyses of classification outputs (see analyses_Classification_performance.Rmd).

Compute principal components.
```{r}
pca <- prcomp(estimations.df.selected[c("huni.chars",
                                       "hrate.chars",
                                       "ttr.chars",
                                       "rm.chars")],
             center = TRUE,
             scale. = TRUE)
print(pca)
summary(pca)
```

Screeplot with loadings of principal components.
```{r}
# compute total variance
variance = pca $sdev^2 / sum(pca $sdev^2)
PCs <- paste("PC", c(1:4), sep = "")
pca.df <- data.frame(PCs, variance)
 
# Scree plot
ggplot(pca.df, aes(x = PCs, y = variance)) + 
  geom_col(alpha = 0.5) +
  geom_point(size = 4)+
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)
```

Bi-plot with ggplot.
```{r}
biplot <- ggbiplot(pca,
              obs.scale = 1,
              var.scale = 1,
              groups = estimations.df.selected$subcorpus,
              ellipse = TRUE,
              circle = FALSE,
              #ellipse.prob = 0.68, # NULL = 0.68
              alpha = 0.3) +
  scale_color_discrete(name = '') +
  theme(legend.direction = 'horizontal',
               legend.position = 'top')
print(biplot)
```

Interpretation: The first two principal components together preserve 94.6% of the original variance in the four variables. PC1 mainly reflects the opposition between repetition rate and entropy (both unigram and rate). Unigram entropy and entropy rate are highly positively correlated (similar vector directions) and have a positive loading on PC1. Modern writing systems fall into the high entropy area, while proto-cuneiform systems tend to have high repetition rates, and low entropies similar to the Aurignacian sequences of signs. TTR is mostly orthogonal to the other three measures and loads mainly on PC2. Aurignacian and Uruk V signs tend to have low TTR, while modern writing systems display a wide range of TTRs (depending on the character inventories).

The normal data ellipse for the Aurignacian sign sequences fully overlaps with the one for the earliest Proto-Cuneiform period (Uruk V), whereas the later two Proto-Cuneiform periods (Uruk IV and Uruk III) start to "move closer" and overlap with the TeDDi (modern writing) ellipse. 

Safe complete figure to file
```{r}
#ggsave("~/Github/UpperPalCombinatorics/figures/pca_biplot.pdf", 
#       biplot, dpi = 300, device = cairo_pdf)
```

Combine figures.
```{r}
combined.plot2 <- ggarrange(length.plot, biplot, ncol = 2, 
                            widths = c(1,2), labels = c("C", "D"))
```

# Final Combination of Plot Panels
Combine figures.
```{r, fig.width = 15, fig.height = 15}
final.plot <- ggarrange(combined.plot1,
                        combined.plot2, ncol = 1, nrow = 2)
final.plot
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/finalPlot_chars.pdf", 
       final.plot, dpi = 300, width = 15, height = 15, device = cairo_pdf)
```

# References 

Jolliffe, I. T., & Cadima, J. (2016). Principal component analysis: a review and recent developments. Philosophical transactions of the royal society A: Mathematical, Physical and Engineering Sciences, 374(2065), 20150202.
