---
title: 'Analyses: Swabian Aurignacian'
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables analyses of feature values for sequences on Swabian Aurgnacian objects (entropy rate, TTR, repetition rate, unigram entropy) as a function of meta information (object type, volume, cave site, etc.). 

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggrepel)
library(ggExtra)
library(gridExtra)
library(ggpubr)
library(ggridges)
library(tidyverse)
library(performance)
library(see)
library(GGally)
library(MASS)
library(forcats)
library(jtools) # for effect plots by predictor variable
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Load Data
Load feature estimations with meta data for the Aurignacian mobile objects.
```{r}
sign.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/signBase_features.csv")
# select Aurignacian data
sign.data <- sign.data[sign.data$archaeological_unit == "Aurignacien", ]
```

# Simple Stats
```{r}
# number of sequences/objects
nrow(sign.data)
# cave sites
table(sign.data$site_name)
# material
table(sign.data$material)
# object_type
table(sign.data$object_type)
```

# Preprocessing
Exclude certain levels of column coding_certainty, and certain numbers of tokens, to check the robustness of results.
```{r}
#sign.data <- sign.data[sign.data$coding_certainty != "Uncertain", ]
#sign.data <- sign.data[sign.data$num.tokens > 10, ]
```

Add columns for object volume and dates.
```{r}
# volume in cm^3 (mm multiplications divided by 1000)
sign.data$volume <- (sign.data$length_mm*sign.data$width_mm*sign.data$depth)/1000
sign.data$date <- as.numeric(substr(sign.data$date_bp_max, 1, 5)) 
```

# Collapse Factors
Collapse some factors of meta information which are uncommon and/or related. Note that using the factor levels as is will lead to problems with model fitting, since there are sometimes very few data points per level.
```{r}
# material: collapse manually
sign.data$material <- fct_collapse(sign.data$material,
             ivory = "ivory",
             antler = "antler",
             bone = "bone",
             other_level = "other")

# object_type: collapse manually
sign.data$object_type <- fct_collapse(sign.data$object_type,
             figurine_anthropomorph = c("figurine anthropomorph",
                          "plaquette"), 
             # note: the "plaquette" is the so-called Adorant with an anthropomorph figure on it.
             figurine_zoomorph = grep('figurine zoomorph.', unique(sign.data$object_type), value = TRUE),
             tool = c("awl", "bâton percé",
                      "rod/bâton", "spatula/lissoir"),
             tube_flute = c("flute", "tube", "tube/flute"),
             personal_ornament = c("personal ornament"),
             other_level = "other")
```

# Simple Statistics Plots
```{r}
# barplot for object counts per cave site
site.barplot <- ggplot(sign.data, aes(x = site_name, fill = site_name)) +
  geom_bar() +
  ylim(0, 200) +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -2.5, hjust = 0.5) +
  geom_text(aes(label = paste(round(after_stat(prop)*100, 1), "%")), 
            stat = "prop", vjust = -1, hjust = 0.5) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Cave Site")
  #coord_flip()
print(site.barplot)

# barplot for object counts per material
material.barplot <- ggplot(sign.data, aes(x = material, fill = material)) +
  geom_bar() +
  ylim(0, 200) +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -2.5, hjust = 0.5) +
  geom_text(aes(label = paste(round(after_stat(prop)*100, 1), "%")), 
            stat = "prop", vjust = -1, hjust = 0.5) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Material")
  #coord_flip()
material.barplot

# barplot for object counts per object_type
type.barplot <- ggplot(sign.data, aes(x = object_type, fill = object_type)) +
  geom_bar() +
  ylim(0, 200) +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -2.5, hjust = 0.5) +
  geom_text(aes(label = paste(round(after_stat(prop)*100, 1), "%")), 
            stat = "prop", vjust = -1, hjust = 0.5) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Object-Type (Carrier)")
  #coord_flip()
type.barplot
```

## Figures Combined
Combine figures.
```{r, fig.width = 12, fig.height = 4, warning = F, message = F}
combined.barplot <- ggarrange(site.barplot, type.barplot, material.barplot,
                           ncol = 3, nrow = 1,
                           labels = "AUTO")
combined.barplot
```

Safe complete figure to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/barPlots_Aurignacian.pdf", 
       combined.barplot, dpi = 300, width = 12, height = 4, device = cairo_pdf)
```

# Sign Stats

## Cave Sites
Get frequency counts for individual sign types per cave site.
```{r}
# initialize dataframe to append results to
caveSigns.df <- data.frame(signs = character(0),
                             Freq = numeric(0), 
                             site_name = character(0))

for (site_name in unique(sign.data$site_name)){
  # get all cleaned codings per cave
  codings <- sign.data[sign.data$site_name == site_name, ]$coding_clean 
  # collapse to one sequence
  codings.collapsed <- paste(codings, collapse = "")
  # remove underscores
  codings.collapsed <- gsub("_", "", codings.collapsed)
  # split into individual signs
  signs <- unlist(strsplit(codings.collapsed, ""))
  # get data frame with frequency counts
  signs.df <- as.data.frame(table(signs))
  # order by frequency
  signs.df <- signs.df[order(-signs.df$Freq), ]
  # add column with site_name
  signs.df$site_name <- rep(site_name, nrow(signs.df))
  # append results to dataframe
  caveSigns.df <- rbind(caveSigns.df, signs.df)
}
```

Barplot for frequency counts per sign type and cave site.
```{r, fig.height = 6, fig.width = 10}
caveSigns.barplot <- ggplot(caveSigns.df, aes(x = reorder(signs, Freq), 
                                          y = Freq, fill = signs)) +
  #ylim(0, 700) +
  geom_text(aes(label = signs), hjust = -1) +
  geom_bar(stat = "identity") +
  facet_wrap(~site_name, ncol = 5) +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Sign Types") +
  ylab("Token Frequency")
caveSigns.barplot
```

## Object Type
Get frequency counts for individual sign types per object type.
```{r}
# initialize dataframe to append results to
objectSigns.df <- data.frame(signs = character(0),
                             Freq = numeric(0), 
                             object_type = character(0))

for (object_type in unique(sign.data$object_type)){
  # get all cleaned codings per object type
  codings <- sign.data[sign.data$object_type == object_type, ]$coding_clean 
  # collapse to one sequence
  codings.collapsed <- paste(codings, collapse = "")
  # remove underscores
  codings.collapsed <- gsub("_", "", codings.collapsed)
  # split into individual signs
  signs <- unlist(strsplit(codings.collapsed, ""))
  # get data frame with frequency counts
  signs.df <- as.data.frame(table(signs))
  # order by frequency
  signs.df <- signs.df[order(-signs.df$Freq), ]
  # add column with site_name
  signs.df$object_type <- rep(object_type, nrow(signs.df))
  # append results to dataframe
  objectSigns.df <- rbind(objectSigns.df, signs.df)
}
```

Barplot for frequency counts per sign type and cave site.
```{r, fig.height = 6, fig.width = 10}
objectSigns.barplot <- ggplot(objectSigns.df, aes(x = reorder(signs, Freq), 
                                          y = Freq, fill = signs)) +
  #ylim(0, 500) +
  geom_text(aes(label = signs), hjust = -1) +
  geom_bar(stat = "identity") +
  facet_wrap(~object_type, ncol = 6) +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Sign Types") +
  ylab("Token Frequency")
objectSigns.barplot
```

## Material
Get frequency counts for individual sign types per material.
```{r}
# initialize dataframe to append results to
materialSigns.df <- data.frame(signs = character(0),
                             Freq = numeric(0), 
                             material = character(0))

for (material in unique(sign.data$material)){
  # get all cleaned codings per cave
  codings <- sign.data[sign.data$material == material, ]$coding_clean 
  # collapse to one sequence
  codings.collapsed <- paste(codings, collapse = "")
  # remove underscores
  codings.collapsed <- gsub("_", "", codings.collapsed)
  # split into individual signs
  signs <- unlist(strsplit(codings.collapsed, ""))
  # get data frame with frequency counts
  signs.df <- as.data.frame(table(signs))
  # order by frequency
  signs.df <- signs.df[order(-signs.df$Freq), ]
  # add column with site_name
  signs.df$material <- rep(material, nrow(signs.df))
  # append results to dataframe
  materialSigns.df <- rbind(materialSigns.df, signs.df)
}
```

Barplot for frequency counts per sign type and cave site.
```{r, fig.height = 6, fig.width = 10}
materialSigns.barplot <- ggplot(materialSigns.df, aes(x = reorder(signs, Freq), 
                                          y = Freq, fill = signs)) +
  #ylim(0, 700) +
  geom_text(aes(label = signs), hjust = -1) +
  geom_bar(stat = "identity") +
  facet_wrap(~ material, ncol = 4) +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Sign Types") +
  ylab("Token Frequency")
materialSigns.barplot
```

## Figures Combined
Combine figures.
```{r, fig.width = 10, fig.height = 18, warning = F, message = F}
signsCombined.barplot <- ggarrange(caveSigns.barplot, objectSigns.barplot, materialSigns.barplot,
                           nrow = 3,
                           labels = "AUTO")
signsCombined.barplot
```

Safe complete figure to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/signsCombinedPlot_Aurignacian.pdf", 
       signsCombined.barplot, dpi = 300, width = 10, height = 14, device = cairo_pdf)
```

## Properties of Objects

## Select Feature
```{r}
# select the statistical feature for further analyses
feature.name <- "hrate.chars"
feature.values <- get(feature.name, sign.data)
```

### Object Type (Carrier)
```{r}
# create plot
objectType.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(object_type, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 3) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Object-Type (Carrier)") +
  theme(legend.position = "none")
objectType.plot
```

### Material
```{r}
# create plot
material.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(material, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 3) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Material") +
  theme(legend.position = "none")
material.plot
```

### Cave Site
```{r}
# create plot
site.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(site_name, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 3) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Cave Site") +
  theme(legend.position = "none")
site.plot
```

### Preservation
```{r}
# create plot
preservation.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(preservation, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 3) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Preservation") +
  theme(legend.position = "none")
preservation.plot
```

### Volume of Object
```{r, warning = F}
# create plot
volume.plot <- ggplot(sign.data, aes(x = feature.values, y = volume, 
                                     size = volume, color = volume)) +
  geom_point(alpha = 0.5) +
  #geom_text_repel(aes(label = object_type), nudge_x = 0.2) +
  geom_smooth(method = "lm") +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Volume (cm^3)") +
  theme(legend.position = "none")
volume.plot
```

### Date
```{r}
# create plot
date.plot <- ggplot(sign.data, aes(x = date, y = feature.values, 
                                   size = volume, color = volume)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(trans = "reverse") +
  #geom_label_repel(aes(label = object_type)) +
  geom_smooth(method = "lm") +
  ylab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  xlab("Date (BP, max)") +
  theme(legend.position = "none")
date.plot
```

## Figures Combined
Combine figures.
```{r, fig.width = 12, fig.height = 7, warning = F, message = F}
combined.plot <- ggarrange(objectType.plot, material.plot, site.plot,
                           preservation.plot, volume.plot, date.plot, 
                           ncol = 3, nrow = 2,
                           labels = "AUTO")
combined.plot
```

Safe complete figure to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_Aurignacian.pdf", 
       combined.plot, dpi = 300, width = 12, height = 7, device = cairo_pdf)
```

# Statistical Model

# Linear Models
We here fit linear modela with feature values (e.g. entropy rates) as response variable, and meta information (e.g. volume, type of object, material etc.) as predictor variables. Some of the predictors are numerical (e.g. volume) while others are categorical (e.g. material, object_type). We also consider interaction terms. Arguably, a linear multiple regression model is appropriate since each object and its respective sequence of signs can be seen as an independent data point. Of course, it is theoretically possible, though very unlikely, that the same paleolithic person has produced a subset of the objects. However, there is no way of knowing this with certainty. Hence, there is no clear evidence for "repeated measures" as in controlled experiments with humans (with subjects and items, cf. Baayen, 2014) -- which would require a mixed-effects structure. 

```{r}
# model with all predictors (no interaction)
linear.model <- lm(feature.values ~ object_type + material + volume + date + preservation + site_name, 
               data = sign.data)
AIC(linear.model) 

# model with interaction between object_type and volume
linear.model.interaction <- lm(feature.values ~ object_type*volume + material + date + preservation + site_name, 
               data = sign.data)
AIC(linear.model.interaction)

# model with interaction between object_type and volume
linear.model.interaction2 <- lm(feature.values ~ object_type + material + date + preservation*volume + site_name, 
               data = sign.data)
AIC(linear.model.interaction2)
```
Interpretation: Adding an interaction term between object type and volume as well as volume and preservation does not improve the model in terms of AIC (a standard criterion is that the AIC would have to decrease by 2 at least.)

## Model Summary
```{r}
summ(linear.model)
```
Interpretation: In the model summary above, the reference level for object_type is "tool". Hence, the entropy rates of sign sequences found on anthropomorph figurines are on average *higher* than the once found on objects classified as tools (borderline significant). A similar effect occurs for zoomorph figurines which also carry more information than tools (significant). Inversely, entropy rates of sequences on tubes and flutes as well as personal ornaments are somewhat *lower* than on tools (significant). Moreover, there is a small positive (but non-siginficant) effect of the volume of an object on the entropy rate, namely, it increases for each cubic centimeter of volume. Younger sequences (c. 30 000 years old) have slightly higher entropy rate than older sequences (c. 40 000 years old), but this effect is non-significant. Lastly, fragmented objects have lower entropy rate than almost complete, and complete once (significant). 

\newpage

## Checking of Model Assumptions
```{r, fig.height = 10, fig.width = 10}
check_model(linear.model)
```

Write to file.
```{r}
png(filename = file.path("~/Github/UpperPalCombinatorics/figures/", "modelCheckingPlot.png"), 
    width = 10, height = 10, units = "in", res = 300)
check_model(linear.model)
dev.off()
```

## Shapiro-Wilk Test for Normality of Residuals
```{r}
shapiro.test(rstandard(linear.model))
```

Interpretation: There is a slight deviation from homogeneity of variance for the residuals (upper right panel), though this is due to just one or two data points deviating for the highest fitted values. There are no overly influential observations, and no collinearity between predictors (middle panels). There is a slight positive skew visible in the distribution of the residuals (lower panels). However, the Shapiro-Wilk test for normality gives a p-value > 0.05, meaning that the distribution does not significantly deviate from the expectation of a normal distribution (null hypothesis).

# Plot Linear Model Coefficients
```{r}
plot_coefs(linear.model, plot.distributions = TRUE, 
           coefs = c("Object type: figurine anthropomorph" = "object_typefigurine_anthropomorph",
                     "Object type: figurine zoomorph" = "object_typefigurine_zoomorph",
                     "Object type: tube or flute" = "object_typetube_flute",
                     "Object type: personal ornament" = "object_typepersonal_ornament",
                     "Material: bone" = "materialbone",
                     "Material: ivory" = "materialivory",
                     "Volume" = "volume",
                     "Age of object" = "date",
                     "Preservation: complete" = "preservationcomplete",
                     "Preservation: fragmented" = "preservationfragmented",
                     "Site name: Hohle Fels" = "site_nameHohle Fels",
                     "Site name: Vogelherd" = "site_nameVogelherd"))
```

# Package References
```{r, echo = F}
print(citation("ggplot2"), style = "text")
print(citation("ggrepel"), style = "text")
print(citation("ggExtra"), style = "text")
print(citation("gridExtra"), style = "text")
print(citation("ggpubr"), style = "text")
print(citation("ggridges"), style = "text")
print(citation("tidyverse"), style = "text")
print(citation("performance"), style = "text")
print(citation("see"), style = "text")
print(citation("MASS"), style = "text")
```

# Literature References
Baayen, R. H. (2008). Analyzing Linguistic Data. A Practical Introduction to Statistics Using R. Cambridge: Cambridge University Press.

Baayen, H. (2014). Multivariate Statistics. In: Podesva, R. J. & Sharma, D. (eds.), Research methods in linguistics. Cambridge: Cambridge University Press, p. 337-373.
