---
title: 'Analyses: Swabian Aurignacian'
author: ""
date: "2/28/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggrepel)
library(ggExtra)
library(gridExtra)
library(ggpubr)
library(ggridges)
library(tidyverse)
```

# Load Data
Load feature estimations with meta data for the Aurignacian mobile objects.
```{r}
sign.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/signBase_features.csv")
# select Aurignacian data
sign.data <- sign.data[sign.data$archaeological_unit == "Aurignacien", ]
```

# Select Feature
```{r}
# select the statistical feature to look at
feature.name <- "huni.chars"
feature.values <- get(feature.name, sign.data)
```

# Properties of Objects

## Object Type (Carrier)
```{r}
# create plot
objectType.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(object_type, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 2) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Object-Type (Carrier)") +
  theme(legend.position = "none")
objectType.plot
```

## Material
```{r}
# create plot
material.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(material, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 2) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Material") +
  theme(legend.position = "none")
material.plot
```

## Cave Site
```{r}
# create plot
site.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(site_name, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 2) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Cave Site") +
  theme(legend.position = "none")
site.plot
```

## Preservation
```{r}
# create plot
preservation.plot <- ggplot(sign.data, aes(x = feature.values, 
                                            y = fct_reorder(preservation, feature.values),
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = feature.name, option = "D") +
  geom_point(shape = 124, alpha = 0.7, size = 2) +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Preservation") +
  theme(legend.position = "none")
preservation.plot
```

## Volume of Object
Create volume column from length, width and depth column.
```{r}
sign.data$volume <- (sign.data$length_mm*sign.data$width_mm*sign.data$depth)/1000
```

```{r}
# create plot
volume.plot <- ggplot(sign.data, aes(x = feature.values, y = volume)) +
  geom_point(alpha = 0.5, size = 2) +
  #geom_label_repel(aes(label = object_type)) +
  geom_smooth(method = "lm") +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Volume (cm^3)")
volume.plot
```

## Date
Create volume column from length, width and depth column.
```{r}
sign.data$date <- as.numeric(substr(sign.data$date_bp_min, 1, 5)) 
```

```{r}
# create plot
date.plot <- ggplot(sign.data, aes(x = feature.values, y = date)) +
  geom_point(alpha = 0.5, size = 2) +
  #geom_label_repel(aes(label = object_type)) +
  geom_smooth(method = "lm") +
  xlab(paste("Feature Values ", "(", feature.name, ")", sep = "")) +
  ylab("Date (BP, min)")
date.plot
```

## Figures Combined
Combine figures.
```{r, fig.width = 16, fig.height = 8}
combined.plot <- grid.arrange(objectType.plot, material.plot, site.plot,
                              preservation.plot, volume.plot,
                              date.plot, ncol = 3)
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_Aurignacian.pdf", 
       combined.plot, dpi = 300, width = 16, height = 8, device = cairo_pdf)
```

