---
title: "Analyses: Feature Value Distributions"
author: ""
date: "23/02/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(ggExtra)
library(gridExtra)
library(ggpubr)
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object (paleolithic signs) and tablet (cuneiform). Comparative sample: A comparative sample of feature values for languages across the world is derived from the Text Data Diversity (TeDDi) sample (Moran et al 2022). 

```{r}
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features_combined.csv")
head(estimations.df)
```

# Preprocessing
```{r}
# select sign strings of certain length
estimations.df <- estimations.df[estimations.df$num.char > 10,]

# select only certain subcorpora for plotting
estimations.df.selected <- estimations.df[estimations.df$subcorpus == "Aurignacien"|
                                          #estimations.df$subcorpus == "Aurignacien/Gravettian"|
                                          estimations.df$subcorpus == "Cuneiform (Uruk V)"|
                                          #estimations.df$subcorpus == "Cuneiform (Uruk IV)"| 
                                          #estimations.df$subcorpus == "Cuneiform (Uruk III)"|  
                                          estimations.df$subcorpus == "TeDDi", ]
```

# Density Plots

Get subset of data points to plot labels.
```{r}
labels.df <- estimations.df.selected[estimations.df.selected$num.char == 11 &
                                       estimations.df.selected$subcorpus != "TeDDi" &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk III)",] 
```

## Entropy rate vs. unigram entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y =  hrate.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  #geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  geom_label_repel(data = labels.df,
                  aes(label = sequence),
    position = position_nudge_repel(x = 0, y = 0.5),
    size = 3,
    max.overlaps = Inf,
    show.legend = FALSE) +
    theme(legend.position = c(.8, .25))
  #facet_wrap(~subcorpus)
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#huni.hrate.chars.plot
```

## TTR vs. repetition rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  theme(legend.position = "none") +
  labs(x = "TTR for characters", y = "Repetition rate for characters") +
  geom_label_repel(data = labels.df,
                  aes(label = sequence),
    position = position_nudge_repel(x = 0.25, y = 0),
    size = 3,
    max.overlaps = Inf,
    show.legend = FALSE)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#ttr.rm.chars.plot
```

Combine figures.
```{r, width = 20, height = 10}
combined.plot <- grid.arrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2)
```

## Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_chars.pdf", 
       combined.plot, dpi = 300, width = 12, height = 6, device = cairo_pdf)
```