---
title: "Analyses: Feature Value Distributions"
author: ""
date: "28/03/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file provides functionality for plotting feature value distributions in density plots. It also contains code for PCA analyses given the values of the four statistical features (entropy rate, unigram entropy, TTR, and repetition rate.)

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
#library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)
library(stringr)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(ggExtra)
library(gridExtra)
library(ggpubr)
library(ggridges)
library(GGally)
```

# Load Data
Load csv file with quantitative feature values for sign sequences per object (paleolithic signs) and tablet (cuneiform). Comparative sample: A comparative sample of feature values for languages across the world is derived from the Text Data Diversity (TeDDi) sample (Moran et al 2022). 

```{r}
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features/features_combined.csv")
head(estimations.df)
```

# Preprocessing
```{r}
# select sign strings of certain length
#estimations.df <- estimations.df[estimations.df$num.char > 1,]

# select only certain subcorpora for plotting
estimations.df.selected <- estimations.df[estimations.df$subcorpus == "Aurignacien"|
                                          #estimations.df$subcorpus == "Aurignacien/Gravettian"|
                                          estimations.df$subcorpus == "Cuneiform (Uruk V)"|
                                          #estimations.df$subcorpus == "Cuneiform (Uruk IV)"| 
                                          estimations.df$subcorpus == "Cuneiform (Uruk III)"|  
                                          estimations.df$subcorpus == "TeDDi", ]
```

# Simple Stats

```{r}
# number of sequences
nrow(estimations.df.selected)
# number of rows (i.e. sequences) per subcorpus
table(estimations.df.selected$subcorpus)
# get mean values of lengths in characters per subcorpus
aggregate(estimations.df.selected$num.char~
            estimations.df.selected$subcorpus, FUN = mean)
# get median values of lengths in characters per subcorpus
aggregate(estimations.df.selected$num.char~
            estimations.df.selected$subcorpus, FUN = median)
```

# Density Plots

## Length
Density plot for lengths in number of characters.
```{r}
length.plot <- ggplot(estimations.df.selected, aes(x = num.char, y = subcorpus, 
                                    fill = subcorpus)) + 
  # add density ridges with quantiles = 2, i.e. median as a line
  geom_density_ridges(alpha = 0.5, quantile_lines = TRUE, quantiles = 2) +
  # limit the x-range 
  xlim(0, 100) +
  xlab("Number of Characters")
length.plot
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/lengthPlot_chars.pdf", 
       length.plot, dpi = 300, width = 5, height = 5, device = cairo_pdf)
```

Get subset of data points to plot labels.
```{r}
# how many data points with a given sequence length?
table(estimations.df.selected$num.char)

# choose subset of data points for labeling
labels.df <- estimations.df.selected[estimations.df.selected$num.char > 9 &
                                       estimations.df.selected$num.char < 11 &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk IV)" &
                                       estimations.df.selected$subcorpus != "Cuneiform (Uruk III)", ]
```

## Entropy rate vs. Unigram Entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(estimations.df.selected, aes(x = huni.chars, y =  hrate.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  xlim(-0.5, 5) +
  ylim(-0.5, 4.5) +
  # geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  geom_label_repel(data = labels.df,
                  aes(label = sequence),
                  size = 3,
                  max.overlaps = Inf,
                  show.legend = FALSE) +
    theme(legend.position = c(.8, .25))
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#huni.hrate.chars.plot
```

## TTR vs. Repetition Rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.selected, aes(x = ttr.chars, y = rm.chars, 
                                    colour = subcorpus, size = num.char)) + 
  geom_point(alpha = 0.3) +
  xlim(-0.1, 1.3) +
  ylim(-0.1, 1.1) +
  theme(legend.position = "none") +
  labs(x = "TTR for characters", y = "Repetition rate for characters") +
  geom_label_repel(data = labels.df,
                  aes(label = sequence),
    position = position_nudge_repel(x = 0.25, y = 0),
    size = 3,
    max.overlaps = Inf,
    show.legend = FALSE)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
#ttr.rm.chars.plot
```

Combine figures.
```{r, fig.width = 20, fig.height = 10}
combined.plot <- grid.arrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2)
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/combinedPlot_chars.pdf", 
       combined.plot, dpi = 300, width = 20, height = 10, device = cairo_pdf)
```

# Correlation Plot
Correlation plots for the statistical feature values.
```{r, fig.width = 12, fig.height = 12}
cor.plot <- ggpairs(estimations.df.selected, columns = c("num.char", "huni.chars", "hrate.chars",
                               "ttr.chars","rm.chars"), 
        aes(color = subcorpus,
            alpha = 0.5))
cor.plot
```

Safe complete figure to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/correlation_plot.pdf", 
       cor.plot, dpi = 300, width = 12, height = 12, device = cairo_pdf)
```


# Principal Component Analysis (PCA)

Compute principal components.
```{r}
pca <- prcomp(estimations.df.selected[c("huni.chars",
                                       "hrate.chars",
                                       "ttr.chars",
                                       "rm.chars")],
             center = TRUE,
             scale. = TRUE)
print(pca)
summary(pca)
```

Screeplot with loadings of principal components.
```{r}
# compute total variance
variance = pca $sdev^2 / sum(pca $sdev^2)
PCs <- paste("PC", c(1:4), sep = "")
pca.df <- data.frame(PCs, variance)
 
# Scree plot
ggplot(pca.df, aes(x = PCs, y = variance)) + 
  geom_col(alpha = 0.5) +
  geom_point(size = 4)+
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)
```

Bi-plot with ggplot.
```{r}
biplot <- ggbiplot(pca,
              obs.scale = 1,
              var.scale = 1,
              groups = estimations.df.selected$subcorpus,
              ellipse = TRUE,
              circle = TRUE,
              #ellipse.prob = 0.68, # NULL = 0.68
              alpha = 0.3) +
  scale_color_discrete(name = '') + 
  theme(legend.direction = 'horizontal',
               legend.position = 'top')
print(biplot)
```

Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/pca_biplot.pdf", 
       biplot, dpi = 300, device = cairo_pdf)
```
