---
title: 'Analyses: Classification Performance'
author: ""
date: "3/1/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(ggpubr)
library(ggridges)
library(tidyverse)
```

# Multi-layer perceptron (MLP)

## Load Data
```{r}
# list files in the performance folders
files.mlp.performance <- list.files(path = "~/Github/UpperPalCombinatorics/results/classifier/MLP/performance", full.names = T)
```

## Preprocessing
```{r}
# initialize empty data frame
performance.mlp.combined <- data.frame(hidden.structure = character(0), 
                                   hidden.size = numeric(0), 
                                   hidden.depth = numeric(0), 
                                   err.fct = character(0), 
                                   act.fct.name = character(0), 
                                   algorithm = character(0), 
                                   baseline = numeric(0), 
                                   accuracy = numeric(0), 
                                   accuracy.pvalue = numeric(0), 
                                   precision = numeric(0), 
                                   recall = numeric(0), 
                                   f1 = numeric(0), 
                                   accuracy.pvalue.corrected = numeric(0), 
                                   comparison = character(0))

# loop through all files and combine them
for (file in files.mlp.performance){
  performance.data <- read.csv(file)
  # apply bonferroni correction
  performance.data$accuracy.pvalue.corrected <- p.adjust(performance.data$accuracy.pvalue, 
                               method = "bonferroni", 
                               n = nrow(performance.data))
  # get the subcorpora involved in the binary classification
  filename.parts <- unlist(strsplit(basename(file), "\\_|\\."))
  subcorpus1 <- filename.parts[6]
  subcorpus2 <- filename.parts[7]
  performance.data$comparison <- rep(paste(subcorpus1, "vs.", subcorpus2), 
                                     nrow(performance.data))
  # add to data frame
  performance.mlp.combined <- rbind(performance.mlp.combined, performance.data)
}
head(performance.mlp.combined)
```

## Density Plots

### F1 Distributions 
Distributions of F1-scores by comparison.
```{r}
f1.plot <- ggplot(performance.mlp.combined, aes(x = f1, 
                                            y = fct_reorder(comparison, accuracy),
                                            fill = after_stat(x))) + 
  geom_density_ridges_gradient(scale = 2, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = "F1-Score", option = "D") +
  geom_point(alpha = 0.2, size = 0.5) +
  xlab("F1-Score") +
  ylab("Comparison") +
  #theme_minimal() +
  theme(legend.position = "none") 
f1.plot
```

## Accuracy Distributions 
Distributions of accuracies by comparison.
```{r}
accuracy.plot <- ggplot(performance.mlp.combined, aes(x = jitter(accuracy, amount = 0.01), 
                                            y = fct_reorder(comparison, accuracy),
                                            fill = after_stat(x))) + 
  geom_density_ridges_gradient(scale = 1, color = "white") +
  scale_fill_viridis_c(name = "Accuracy", option = "D") +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_point(shape = 124, size = 4, aes(x = baseline, y = comparison)) +
  xlab("Accuracy") +
  ylab("Comparison") +
  theme(legend.position = "none") 
accuracy.plot
```

## P-value Distributions 
Distributions of p-values by comparison.
```{r}
pvalue.plot <- ggplot(performance.mlp.combined, aes(x = jitter(accuracy.pvalue.corrected, amount = 0.05), 
                                            y = fct_reorder(comparison, accuracy),
                                            fill = after_stat(x))) + 
  geom_density_ridges_gradient(scale = 1, color = "white") +
  scale_fill_viridis_c(name = "p-values", option = "D") +
  geom_point(alpha = 0.2, size = 0.5) +
  xlab("p-values") +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank())
  #geom_density(alpha = 0.5)
pvalue.plot
```

## Combine Plots
```{r, width = 10, height = 5}
combined.plot <- grid.arrange(accuracy.plot, pvalue.plot, ncol = 2, widths = c(1.8,1))
```

Safe to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/performancePlot_MLP.pdf", 
       combined.plot, dpi = 300, width = 10, height = 5, device = cairo_pdf)
```

# K-Nearest-Neighbors (KNN)

## Load Data
```{r}
# list files in the performance folders
files.knn.performance <- list.files(path = "~/Github/UpperPalCombinatorics/results/classifier/KNN/performance", full.names = T)
```

## Preprocessing
```{r}
# initialize empty data frame
performance.knn.combined <- data.frame(k = numeric(0), 
                                   baseline = numeric(0), 
                                   accuracy = numeric(0), 
                                   accuracy.pvalue = numeric(0), 
                                   precision = numeric(0), 
                                   recall = numeric(0), 
                                   f1 = numeric(0), 
                                   accuracy.pvalue.corrected = numeric(0), 
                                   comparison = character(0))

# loop through all files and combine them
for (file in files.knn.performance){
  performance.data <- read.csv(file)
  # apply bonferroni correction
  performance.data$accuracy.pvalue.corrected <- p.adjust(performance.data$accuracy.pvalue, 
                               method = "bonferroni", 
                               n = nrow(performance.data))
  # get the subcorpora involved in the binary classification
  filename.parts <- unlist(strsplit(basename(file), "\\_|\\."))
  subcorpus1 <- filename.parts[3]
  subcorpus2 <- filename.parts[4]
  performance.data$comparison <- rep(paste(subcorpus1, "vs.", subcorpus2), 
                                     nrow(performance.data))
  # add to data frame
  performance.knn.combined <- rbind(performance.knn.combined, performance.data)
}
head(performance.knn.combined)
```

## Density Plots

### F1 Distributions 
Distributions of F1-scores by comparison.
```{r}
f1.plot <- ggplot(performance.knn.combined, aes(x = jitter(f1, amount = 0.05), 
                                            y = fct_reorder(comparison, accuracy), 
                                            fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = "F1-Score", option = "C") +
  xlab("F1-Score") +
  ylab("Comparison") +
  theme(legend.position = "none")
  #geom_density(alpha = 0.5)
f1.plot
```

### Accuracy Distributions 
Distributions of accuracies by comparison.
```{r}
accuracy.plot <- ggplot(performance.knn.combined, aes(x = jitter(accuracy, amount = 0.05), 
                                                  y = fct_reorder(comparison, accuracy), 
                                                  fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = "Accuracy", option = "D") +
  geom_point(shape = 124, size = 3, aes(x = baseline, y = comparison)) +
  xlab("Accuracy") +
  ylab("Comparison") +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank())
accuracy.plot
```

### P-value Distributions 
Distributions of p-values by comparison.
```{r}
pvalue.plot <- ggplot(performance.knn.combined, aes(x = jitter(accuracy.pvalue.corrected, amount = 0.05),
                                                y = fct_reorder(comparison, accuracy),
                                                fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 1, color = "white",
                               quantile_lines = T, quantiles = 2) +
  scale_fill_viridis_c(name = "P-value", option = "A") +
  xlab("p-value") +
  ylab("Comparison") +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank())
  #geom_density(alpha = 0.5)
pvalue.plot
```

## Combine Plots
```{r, width = 10, height = 5}
combined.plot <- grid.arrange(f1.plot, accuracy.plot, pvalue.plot, ncol = 3, widths = c(2,1,1))
```

Safe to file.
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/performancePlot_KNN.pdf", 
       combined.plot, dpi = 300, width = 10, height = 5, device = cairo_pdf)
```
