---
title: 'Analyses: Classification with KNN'
author: ""
date: "01/03/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
k-nearest neighbor analyses of the feature vectors per character sequence (loaded from ~/results/features_combined.csv). The results are stored in ~/results/classifier/.

## Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r, message = F}
library(dplyr)
library(class)
library(gmodels)
library(caret)
```

# Load Data
Load data table with estimated feature values per sequence. 
```{r}
# load feature value estimations 
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features/features_combined.csv")
head(estimations.df)
```

Select subcorpora for binary classification.
```{r}
# choose subcorpora to compare 
unique(estimations.df$subcorpus)
subcorpus1 <- "Cuneiform (Uruk V)"
subcorpus2 <- "Cuneiform (Uruk III)"
selected <- c(subcorpus1, subcorpus2)
estimations.subset <- estimations.df[estimations.df$subcorpus %in% selected, ]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
# first column is subcorpus column, the others are the features
# be aware that this needs to be changed in case of other table format
estimations.scaled <- cbind(estimations.subset[, c("subcorpus", "sequence", "num.char")], 
                            scale(estimations.subset[, c("huni.chars", "hrate.chars", 
                                                         "ttr.chars", "rm.chars")]))
```
# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test samples with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.67, 0.33))

# Generate training set (select rows randomly by using datasample vector)
train <- estimations.scaled[datasample == 1, 1:ncol(estimations.scaled)]
# Generate test set 
test <- estimations.scaled[datasample == 2, 1:ncol(estimations.scaled)]
```

# Initialize data frame
```{r}
knn.results <- data.frame(k = numeric(0), accuracy = numeric(0), precision = numeric(0), 
                             recall = numeric(0), f1 = numeric(0))
```

# Building knn classifier
```{r}
# choose maximum number of neighbors n
n = 10
# run a loop over different numbers of neighbors up to n
for (k in 1:n){
  # knn estimation of labels
  predictions.knn <- knn(train = train[, c("huni.chars", 
                                           "hrate.chars", 
                                            "ttr.chars", 
                                           "rm.chars")],
                         test = test[, c("huni.chars", 
                                           "hrate.chars", 
                                            "ttr.chars", 
                                           "rm.chars")], 
                         cl = train[, "subcorpus"], k = k) 
  
  # model evaluation
  # creating a dataframe from known (true) test labels
  test.labels <- test[, "subcorpus"]
  # combining predicted and known classes
  class.comparison <- data.frame(predictions.knn, test.labels)
  # giving appropriate column names
  names(class.comparison) <- c("predicted", "observed")
  # inspecting our results table
  head(class.comparison)
  # get confusion matrix
  cm <- confusionMatrix(data = class.comparison$predicted, 
                      reference = as.factor(class.comparison$observed))
  print(cm)
  # get precision, recall, and f1 from the output list of confusionMatrix()
  accuracy <- cm$overall['Accuracy']
  f1 <- cm[["byClass"]]["F1"]
  recall <- cm[["byClass"]]["Recall"]
  precision <- cm[["byClass"]]["Precision"]
  
  # prepare data frame with results
  local.results <- data.frame(k, accuracy, precision, recall, f1, row.names = NULL)
  local.results.rounded <- round(local.results, 2)
  # print(local.results.rounded)
  knn.results <- rbind(knn.results, local.results.rounded)
  # give predictions alongside the other information in data frame
  knn.predictions.df <- cbind(test, predictions.knn)
}
print(knn.results)
```

Write to file. 
```{r}
# write results
write.csv(knn.results, file = paste(c("~/Github/UpperPalCombinatorics/results/classifier/KNN/results_KNN_",
                                              subcorpus1, "_",
                                              subcorpus2, "_",
                                              ".csv"
                                              ), collapse = ""),
                                      row.names = F)

# write df with predictions for post-hoc analyses
# note that this writes the predictions only for the last run of the loop, i.e. the highest k
write.csv(knn.predictions.df, file = paste(c("~/Github/UpperPalCombinatorics/results/classifier/KNN/predictions_KNN_",
                                              subcorpus1, "_",
                                              subcorpus2, "_",
                                              ".csv"
                                              ), collapse = ""),
                                      row.names = F)
```
