---
title: 'Analyses: Classification with MLP'
author: ""
date: "03/01/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

# Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
# install the latest version of neuralnet with bug fixes: devtools::install_github("bips-hb/neuralnet")
library(neuralnet)
library(sigmoid)
library(caret)
```

# Load Data
Load data table with estimated feature values per sequence. 
```{r}
# load feature value estimations 
estimations.df <- read.csv("~/Github/UpperPalCombinatorics/results/features/features_combined.csv")
head(estimations.df)
```

Choose subcorpora for binary classification.
```{r}
# choose subcorpora to compare: "Aurignacien", "TeDDi", "Cuneiform (Uruk V)",
# "Cuneiform (Uruk IV)", "Cuneiform (Uruk III)".
subcorpus1 <- "Cuneiform (Uruk IV)"
subcorpus2 <- "Cuneiform (Uruk III)"
selected <- c(subcorpus1, subcorpus2)
estimations.subset <- estimations.df[estimations.df$subcorpus %in% selected, ]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
# first three columns are meta information, the others are the feature values
# be aware that this needs to be changed in case of other table format
estimations.scaled <- cbind(estimations.subset[, c("subcorpus", "sequence", "num.char")], 
                            scale(estimations.subset[, c("huni.chars", "hrate.chars", 
                                                         "ttr.chars", "rm.chars")]))
```

# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test samples with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.67, 0.33))
# Generate training set (select rows randomly by using datasample vector)
train <- estimations.scaled[datasample == 1, 1:ncol(estimations.scaled)]
# Generate test set 
test <- estimations.scaled[datasample == 2, 1:ncol(estimations.scaled)]
# number of data points in test set
test.n <- nrow(test)
```

# Implement MLP classifier
This is partly based on code given at http://uc-r.github.io/ann_classification (last accessed 18.01.2023). Theactivation function (tanh), error function (SSE), and backpropagation algorithm (rprop+) are here chosen according to the best model performance for sequences of 100 characters reported in Bentz (2023).
```{r, message = F, warning = F}
# Generate list of hidden unit architectures (number of layers and units)
# Choose number of random architectures
arch.no <- 100
# initialize empty list of hidden unit architectures
hidden.list <- vector(mode = "list", length = arch.no)
# set seed for reproducible results
set.seed(09012020)
# run loop to generate number of layers and hidden units randomly
for (i in 1:arch.no) {
  layer.no <- round(runif(1, min = 1, max = 4), 0)
  hidden.units <- round(runif(layer.no, min = 1, max = 5), 0)
  hidden.list[[i]] <- hidden.units
}

# Manual list of architectures for testing 
# hidden.list <- list(c(4,4))

# initialize dataframe to append results to
results.df <- data.frame(hidden.structure = character(0), 
                         hidden.size = numeric(0), hidden.depth = numeric(0),
                         err.fct = character(0), act.fct.name = character(0), 
                         algorithm = character(0), baseline = numeric(0),
                         accuracy = numeric(0), accuracy.pvalue = numeric(0), 
                         precision = numeric(0), recall = numeric(0), 
                         f1 = numeric(0), test.n = numeric(0) 
                         ) 

# set random seed for reproducibility
set.seed(123)
# counter 
counter <- 0
# start time
start_time <- Sys.time()
for (hidden in hidden.list) {
  # Training
  # choose error function
  err.fct = 'sse' # options: 'sse', 'ce'
  # choose activation function
  act.fct = 'tanh' # options: 'logistic', 'tanh', softplus, relu, 
  # beware: the last two options have to be written without quotes ''
  # note: softplus, relu, and tanh only work properly with sse as err.fct
  act.fct.name = 'tanh' # make sure to give name as character string (for later storage)
  # choose algorithm 
  algorithm = 'rprop+' # options: 'rprop+', 'rprop-', 'backprop', 'sag', 'slr' 
  # train classifier with neuralnet() function
  try({ # if the processing fails for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running 
  classifier.mlp <- neuralnet(subcorpus == subcorpus1 ~ .,
                     data = train[, c("subcorpus",
                                      "huni.chars", 
                                      "hrate.chars", 
                                      "ttr.chars", 
                                      "rm.chars")],
                     hidden = hidden, # hidden layer architecture
                     threshold = 0.1, # defaults to 0.01
                     rep = 1, # number of reps in which new initial values are used,
                     # (essentially the same as a for loop)
                     stepmax = 100000, # defaults to 100K
                     linear.output = FALSE,
                     algorithm = algorithm, # defaults to "rprop+",
                     # i.e. resilient backpropagation
                     # learningrate = 0.1,
                     err.fct = err.fct,
                     act.fct = act.fct,
                     likelihood = TRUE,
                     lifesign = 'minimal')
  # classifier.mlp
  # results matrix (each column represents one repetition)
  # classifier.mlp$result.matrix
  
  # Prediction
  # Get prediction using the predict() function
  mlp.predictions <- predict(classifier.mlp, test, 
                           rep = which.min(classifier.mlp$result.matrix[1,]), # Predict response values 
                           # based on the "best" repetition (epoche), i.e. the one with the lowest error.
                           all.units = FALSE)
  # assign a label according to the rule that the label is the label of subcorpus1 
  # if the prediction probability is >0.5, else assign label of subcorpus2. 
  mlp.predictions.rd <- ifelse(mlp.predictions > 0.5, subcorpus1, subcorpus2)
  
  # Model Evaluation
  # creating a dataframe from known (true) test labels
  test.labels <- data.frame(test$subcorpus)
  # combining predicted and known classes
  class.comparison <- data.frame(mlp.predictions.rd, test.labels)
  # giving appropriate column names
  names(class.comparison) <- c("predicted", "observed")
  # inspecting our results table
  head(class.comparison)
  # get confusion matrix
  cm <- confusionMatrix(as.factor(class.comparison$predicted), 
                      reference = as.factor(class.comparison$observed))
  #print(cm)
  # get performance metrics from the output list of confusionMatrix()
  baseline <- cm$overall['AccuracyNull']
  accuracy <- cm$overall['Accuracy']
  accuracy.pvalue <- cm$overall['AccuracyPValue']
  f1 <- cm[["byClass"]]["F1"]
  recall <- cm[["byClass"]]["Recall"]
  precision <- cm[["byClass"]]["Precision"]
  
  # unname and round resulting values (except for pvalue, 
  # since this needs to be corrected later)
  baseline <- round(unname(baseline), 2)
  accuracy <- round(unname(accuracy), 2)
  f1 <- round(unname(f1), 2)
  recall <- round(unname(recall), 2)
  precision <- round(unname(precision), 2)

  # append results to dataframe
  hidden.structure <- paste(unlist(hidden), collapse = ",")
  hidden.size <- sum(unlist(hidden))
  hidden.depth <- length(hidden)
  local.df <- data.frame(hidden.structure, hidden.size, 
                         hidden.depth, err.fct, act.fct.name, algorithm, 
                         baseline, accuracy, accuracy.pvalue,
                         precision, recall, f1, test.n)
  results.df <- rbind(results.df, local.df)
  })
  counter <- counter + 1
  print(counter)
}
end_time <- Sys.time()
proc.time <- end_time - start_time
print(proc.time)
```

Write performance metrics to file.
```{r}
# classification results summaries 
write.csv(results.df, file = paste(c("~/Github/UpperPalCombinatorics/results/classifier/MLP/performance/performance_MLP_",
                                              subcorpus1, "_",
                                              subcorpus2,
                                              ".csv"
                                              ), collapse = ""),
                                      row.names = F)
```

Safe label predictions alongside the other columns in the test data set. Note that these are the results for the *last run* of the for-loop.
```{r}
mlp.predictions.df <- cbind(test, mlp.predictions.rd)
```

```{r}
# safe predictions 
write.csv(mlp.predictions.df, file = paste(c("~/Github/UpperPalCombinatorics/results/classifier/MLP/predictions/predictions_MLP_",
                                              subcorpus1, "_",
                                              subcorpus2,
                                              ".csv"
                                              ), collapse = ""),
                                      row.names = F)
```

# Visualize the NN
Visualize the NN with the best weights after training. Note this is the *last run* of the for loop through different architectures.
```{r}
#mlp.plot <- plot(classifier.mlp, rep = 'best', show.weights = F)
#mlp.plot
```

# References
Bentz, Christian (2023). The Zipfian Challenge: Learning the statistical fingerprint of natural
languages. CoNLL, Singapore. https://aclanthology.org/2023.conll-1.3/
