---
title: 'Analyses: Modern Writing (TeDDi)'
author: "Chris Bentz"
date: "2/28/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(ggplot2)
library(ggrepel)
library(ggExtra)
library(gridExtra)
library(ggpubr)
```

# Load Data
Load feature estimations with meta data for the Aurignacian mobile objects.
```{r}
teddi.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/teddi_features.csv")
```

# Preselection
```{r}
teddi.data <- teddi.data[teddi.data$num.char > 20 &
                           teddi.data$num.char < 200, ]
# select particular writing systems
# unique(teddi.data$writing_system)
teddi.data <- teddi.data[teddi.data$writing_system == "Arab"|
                           teddi.data$writing_system == "Geor"| 
                           teddi.data$writing_system == "Hans"|
                           teddi.data$writing_system == "Thai", ]
```

# TeDDi Specific Plots
Comparisons of different writing systems in TeDDi.

## Entropy rate vs. unigram entropy
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
huni.hrate.chars.plot <- ggplot(teddi.data, aes(x = huni.chars, y = hrate.chars, 
                                    colour = writing_system, size = num.char)) + 
  geom_point(alpha = 0.3) +
  #geom_density2d(alpha = .8) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  theme(legend.position = c(.8,.2))
  #facet_wrap(~subcorpus)
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
huni.hrate.chars.plot
```

## TTR vs. repetition rate
```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ttr.rm.chars.plot <- ggplot(teddi.data, aes(x = ttr.chars, y = rm.chars, 
                                    colour = writing_system, size = num.char)) + 
  geom_point(alpha = 0.3) +
  #geom_density2d(alpha = .5) +
  labs(x = "TTR for characters", y = "Repetition rate for characters") +
  theme(legend.position = "none")
  #facet_wrap(~subcorpus)
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, groupColour = T, 
                                  type = "density")
ttr.rm.chars.plot
```

Combine figures.
```{r}
combined.plot <- grid.arrange(huni.hrate.chars.plot, ttr.rm.chars.plot, ncol = 2)
```

## Safe complete figure to file
```{r}
ggsave("~/Github/UpperPalCombinatorics/figures/writingSystem_plot.pdf", 
       combined.plot, dpi = 300, width = 12, height = 6, device = cairo_pdf)
```

