---
title: "Stabilization Analyses"
author: ""
date: "11/03/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Description
This file enables stabilization analyses for the feature value estimations. In other words, feature values are not estimated for a single, pre-defined number of characters, but are estimated for discrete steps (defined with "stepsize"), and for a given maximum number of characters (defined as ``n'').

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(plyr)
library(entropy)
library(ggExtra)
library(gsubfn)
# library(devtools)
# install_github("dimalik/Hrate")
library(Hrate)
```

## Load Data
Read the preprocessed files with character strings per row. 
```{r}
# give file paths to the files to be processed
teddi <- read.csv("~/Github/UpperPalCombinatorics/data/TeDDi/final_sample.csv")
signbase <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/signBase_randomized.csv")
cdli.urukV <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukV/signs_clean.csv")
cdli.urukIV <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukIV/signs_clean.csv")
cdli.urukIII <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukIII/signs_clean.csv")
```

# Stabilization analysis per file
```{r}
# set counter
counter = 0
# set the maximal number of units (n), and the stepsize for stabilization analysis
# (i.e. in steps of how many units are values calculated?)
n = 100
stepsize = 10
# initialize dataframe to append results to
stabilization.df <- data.frame(filename = character(0), subcorpus = character(0), 
                               code = character(0), huni.chars = numeric (0), 
                               hrate.chars = numeric(0), ttr.chars = numeric(0), 
                               rm.chars = numeric(0), units = numeric(0)) 
# start time
start_time <- Sys.time()
for (file in file.list) 
{
  try({ # if the processing failes for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running  
  
  # basic processing
  # loading textfile
  textfile <- scan(file, what = "char", quote = "", 
                   comment.char = "", encoding = "UTF-8", sep = "\n" , skip = 7, nmax = 20) 
  # skip 7 first lines, nmax gives the maximum number of lines to be read,
  # note that reading more lines will considerably increase processing time.
  # remove annotations marked by '<>'
  textfile <- gsub("<.*>","",textfile) 
  # print(head(textfile))
  # get filename
  filename <- basename(file) 
  #print(filename) # for visual inspection
  # get subcorpus category
  subcorpus <- sub("_.*", "", filename)
  # print(subcorpus) # for visual inspection
  # get the three letter identification code + the running number
  code <- substring(substring(filename, regexpr("_", filename) + 1), 1, 8)
  
  # Split into individual characters/signs 
  # remove tabs and parentheses, as well as star signs `*' and plus signs `+´
  # note that this might have to be tuned according to the text files included
  textfile <- str_replace_all(textfile, c("\\\t" = "", "\\(" = "", "\\)" = "", 
                                       "\\]" = "", "\\[" = "",  "\\}" = "", 
                                       "\\{" = "", "\\*" = "", "\\+" = ""))
  # split the textfile into individual utf-8 characters. Note that white spaces are
  # counted as utf-8 characters here.
  chars <- unlist(strsplit(textfile, ""))
  chars <- chars[1:n] # use only maximally n units
  chars <- chars[!is.na(chars)] # remove NAs for vectors which are already shorter
  # than n
  # chars <- chars[chars != " "] # remove white spaces from character vector
  # use "next" statement to exclude files with less than x characters
  if (length(chars) < 100) {
    next
  }
  # run loop with stepsizes
  # define the number of units (i.e. characters) used for analyses (note that k is      
  # always either equal to or smaller than n)
  k = length(chars)
  for (i in 1:(k/stepsize))
  {  
    # unigram entropy estimation
    # define substring of chars vector by stepsize
    sub.chars <- chars[1:(i*stepsize)]
    # calculate unigram entropy for characters
    chars.df <- as.data.frame(table(sub.chars))
    # print(chars.df)
    huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
    # entropy rate estimation
    # note: the values chosen for max.length and every.word will crucially 
    # impact processing time. max.length = NULL means all units in the file are     
    # considered.
    hrate.chars <- get.estimate(text = sub.chars, every.word = 1, 
                                max.length = NULL)  
    # calculate type-token ratio (ttr)
    ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
    
    # calculate repetition measure according to Sproat (2014)
    # the overall number of repetitions is the sum of frequency counts minus 1. 
    R <- sum(chars.df$Freq)-1 
    # calculate the number of adjacent repetitions
    r = 0
    if (length(sub.chars) > 1){
      for (j in 1:(length(sub.chars)-1)){
        if (sub.chars[j] == sub.chars[j+1]){
          r = r + 1
        } else {
          r = r + 0
        }
      }
      # calculate the repetition measure
      rm.chars <- r/R
    } else {
      rm.chars <- "NA"  
    }
    
    # append results to dataframe
    local.df <- data.frame(filename, subcorpus, code, huni.chars, hrate.chars, 
                           ttr.chars, rm.chars, units = i*stepsize)
    stabilization.df <- rbind(stabilization.df, local.df)
  }
  # counter
  counter <- counter + 1
  # print(counter)
  })
}
end_time <- Sys.time()
end_time - start_time
head(stabilization.df)
```

# Stabilization plots

## Unigram entropy characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
huni.chars.plot <- ggplot(stabilization.df, aes(x = units, y = huni.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "unigram entropy for characters") +
  facet_wrap(~code)
huni.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_huni_chars.pdf", huni.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## Entropy rate characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
hrate.chars.plot <- ggplot(stabilization.df, aes(x = units, y = hrate.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "entropy rate for characters") + 
  facet_wrap(~code)
hrate.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_hrate_chars.pdf", hrate.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## TTR characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
ttr.chars.plot <- ggplot(stabilization.df, aes(x = units, y = ttr.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "TTR for characters") +
  facet_wrap(~code)
ttr.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_ttr_chars.pdf", ttr.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## Repetition rate characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
rm.chars.plot <- ggplot(stabilization.df, aes(x = units, y = rm.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "RM for characters") +
  facet_wrap(~code)
rm.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_rm_chars.pdf", rm.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```