---
title: "Stabilization Analyses"
author: ""
date: "12/03/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Description
This file enables stabilization analyses for the feature value estimations. In other words, feature values are not estimated for a single, pre-defined number of characters, but are estimated for discrete steps (defined with "stepsize"), and for a given maximum number of characters (defined as ``n'').

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(plyr)
library(entropy)
library(ggExtra)
library(gsubfn)
# library(devtools)
# install_github("dimalik/Hrate")
library(Hrate)
```

# Load Data
Read the preprocessed files with character strings per row. 
```{r}
# give file paths to the files to be processed
teddi <- read.csv("~/Github/UpperPalCombinatorics/data/TeDDi/final_sample.csv")
signbase <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/signBase_randomized.csv")
cdli.urukV <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukV/signs_clean.csv")
cdli.urukIV <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukIV/signs_clean.csv")
cdli.urukIII <- read.csv("~/Github/UpperPalCombinatorics/data/CDLI/UrukIII/signs_clean.csv")
```

# Preprocessing
```{r}
# TeDDi
# select relevant columns
teddi.selected <- teddi[c("filename", "text","writing_system")]
# rename columns
colnames(teddi.selected) <- c("id", "sequence", "subcorpus")
# add column which identifies the corpus
teddi.selected$corpus <- rep("TeDDi", nrow(teddi.selected))

# SignBase
signbase <- signbase[signbase$archaeological_unit == "Aurignacien", ]
signbase.selected <- signbase[c("object_id", "archaeological_unit", "site_name", "coding_clean")]
colnames(signbase.selected) <- c("id", "corpus", "subcorpus", "sequence")

# Proto-Cuneiform
## Uruk V
cdli.urukV.selected <- cdli.urukV[c("tablet.id", "signs.expand.vec")]
colnames(cdli.urukV.selected) <- c("id", "sequence")
cdli.urukV.selected$corpus <- rep("Proto-Cuneiform", nrow(cdli.urukV.selected))
cdli.urukV.selected$subcorpus <- rep("Uruk V", nrow(cdli.urukV.selected))
## Uruk IV
cdli.urukIV.selected <- cdli.urukIV[c("tablet.id", "signs.expand.vec")]
colnames(cdli.urukIV.selected) <- c("id", "sequence")
cdli.urukIV.selected$corpus <- rep("Proto-Cuneiform", nrow(cdli.urukIV.selected))
cdli.urukIV.selected$subcorpus <- rep("Uruk IV", nrow(cdli.urukIV.selected))
## Uruk III
cdli.urukIII.selected <- cdli.urukIII[c("tablet.id", "signs.expand.vec")]
colnames(cdli.urukIII.selected) <- c("id", "sequence")
cdli.urukIII.selected$corpus <- rep("Proto-Cuneiform", nrow(cdli.urukIII.selected))
cdli.urukIII.selected$subcorpus <- rep("Uruk III", nrow(cdli.urukIII.selected))

# Combined Data Frame
data.combined <- rbind(teddi.selected, signbase.selected, cdli.urukV.selected,
                       cdli.urukIV.selected, cdli.urukIII.selected)
```

# Stabilization Analysis per Sequence
```{r}
# set counter
counter = 0
# set the maximal number of characters (max.n), the minimum number (min.n),
# and the stepsize for stabilization analysis 
# (i.e. in steps of how many units are values calculated?).
# Note: it should be the case that min.n >= stepsize
max.n = 200
min.n = 100
stepsize = 10
# initialize dataframe to append results to
stabilization.df <- data.frame(id = character(0), corpus = character(0), 
                               subcorpus = character(0), sequence = character(0),
                               huni.chars = numeric (0), hrate.chars = numeric(0), 
                               ttr.chars = numeric(0), rm.chars = numeric(0), 
                               num.char = numeric(0)) 
# start time
start_time <- Sys.time()
for (i in 1:nrow(data.combined)) 
{
  try({ # if the processing fails for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running  
  
  # get meta information per row
  row.data <- data.combined[i, ]
  id <- row.data$id
  corpus <- row.data$corpus
  subcorpus <- row.data$subcorpus
  sequence <- row.data$sequence
  num.char <- nchar(row.data$sequence)
  
  # if/else statement to distinguish between the proto-cuneiform signs
  # and individial UTF-8 characters in the other sequences
  if (corpus == "Proto-Cuneiform"){
    chars <- unlist(strsplit(sequence, " "))
  } else {
    chars <- unlist(strsplit(sequence, ""))
  }
   
  # preprocess text with bag-of-words normalization
  # note: this is an important step! Without it, entropy rate estimation will fail
  # to yield realistic results due to issues with processing special characters.
  chars <- PreprocessText(chars)
  
  # maximal length in chars
  chars <- chars[1:max.n] # use only maximally n units
  # remove NAs for vectors which are already shorter than max.n
  chars <- chars[!is.na(chars)] 
  
  # use "next" statement to exclude files with less than x characters
  if (length(chars) < min.n) {
    next
  }
  
  # run loop with stepsizes
  # define the number of units (i.e. characters) used for analyses (note that k is      
  # always either equal to or smaller than n)
  k = length(chars)
  for (j in 1:(k/stepsize))
  {  
    # unigram entropy estimation
    # define substring of chars vector by stepsize
    sub.chars <- chars[1:(j*stepsize)]
    # calculate unigram entropy for characters
    chars.df <- as.data.frame(table(sub.chars))
    # print(chars.df)
    huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
    # entropy rate estimation
    # note: the values chosen for max.length and every.word will crucially 
    # impact processing time. max.length = NULL means all units in the file are     
    # considered.
    hrate.chars <- get.estimate(text = sub.chars, every.word = 1, 
                                max.length = NULL)  
    # calculate type-token ratio (ttr)
    ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
    
    # calculate repetition measure according to Sproat (2014)
    # the overall number of repetitions is the sum of frequency counts minus 1. 
    R <- sum(chars.df$Freq)-1 
    # calculate the number of adjacent repetitions
    r = 0
    if (length(sub.chars) > 1){
      for (m in 1:(length(sub.chars)-1)){
        if (sub.chars[m] == sub.chars[m+1]){
          r = r + 1
        } else {
          r = r + 0
        }
      }
      # calculate the repetition measure
      rm.chars <- r/R
    } else {
      rm.chars <- "NA"  
    }
    
    # append results to dataframe
    local.df <- data.frame(id, corpus, subcorpus, sequence,
                           huni.chars, hrate.chars, 
                           ttr.chars, rm.chars, num.char = j*stepsize)
    stabilization.df <- rbind(stabilization.df, local.df)
  }
  # counter
  counter <- counter + 1
  print(counter)
  })
}
end_time <- Sys.time()
end_time - start_time
head(stabilization.df)
```

# Stabilization plots

## Unigram entropy characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
huni.chars.plot <- ggplot(stabilization.df, aes(x = num.char, y = huni.chars,
                                                color = subcorpus)) + 
  geom_point(alpha = 0.3, size  = 2) +
  geom_smooth() +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "unigram entropy for characters") +
  facet_wrap(~id)
huni.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/UpperPalCombinatorics/figures/stabilization_huni_chars.pdf", huni.chars.plot, dpi = 300, 
       width = 20, height = 20, 
       scale = 1, device = cairo_pdf)
```

## Entropy rate characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
hrate.chars.plot <- ggplot(stabilization.df, aes(x = num.char, y = hrate.chars,
                                                colour = corpus)) + 
  geom_point(alpha = 0.3, size  = 2) +
  geom_smooth() +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "entropy rate for characters") + 
  facet_wrap(~id)
hrate.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/UpperPalCombinatorics/figures/stabilization_hrate_chars.pdf", hrate.chars.plot, dpi = 300, 
       width = 20, height = 20, 
       scale = 1, device = cairo_pdf)
```

## TTR characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
ttr.chars.plot <- ggplot(stabilization.df, aes(x = num.char, y = ttr.chars,
                                                colour = corpus)) + 
  geom_point(alpha = 0.3, size  = 2) +
  geom_smooth() +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "TTR for characters") +
  facet_wrap(~id)
ttr.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/UpperPalCombinatorics/figures/stabilization_ttr_chars.pdf", ttr.chars.plot, dpi = 300, 
       width = 20, height = 20, 
       scale = 1, device = cairo_pdf)
```

## Repetition rate characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
rm.chars.plot <- ggplot(stabilization.df, aes(x = num.char, y = rm.chars,
                                                colour = corpus)) + 
  geom_point(alpha = 0.3, size  = 2) +
  geom_smooth() +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "RM for characters") +
  facet_wrap(~id)
rm.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/UpperPalCombinatorics/figures/stabilization_rm_chars.pdf", rm.chars.plot, dpi = 300,
       width = 20, height = 20, 
       scale = 1, device = cairo_pdf)
```