---
title: "Preprocessing for SignBase Sequences"
author: ''
date: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables preprocessing of Upper-Paleolithic sign sequences as published at https://www.signbase.org. The csv file with latest objects and sign codings can be downloaded from there. The file used for the analyses here is also given in the github repo under data/signBase/.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Preprocessing
Preprocessing and storage of cleaned file for signs. Load data from github repo and remove missing rows.
```{r}
# load data from github repo
sign.data <- read.csv(
  "~/Github/UpperPalCombinatorics/data/signBase/Sb2Artifact-2024-04-26.csv")
# remove certain columns
sign.data <- subset(sign.data, select = -c(short_description, references))
# remove rows for which there is no sign coding
sign.data <- sign.data[!sign.data$sign_coding == "", ]
# remove rows with "very uncertain" sign codings
sign.data <- sign.data[!sign.data$coding_certainty == "Very uncertain", ]
```

Remove UTF-8 characters which do not represent actual signs on the objects, i.e annotations. A table with these distinctions is also provided at www.signbase.org. 
```{r}
# remove some special characters 
coding.clean <- gsub('[!?]', '', sign.data$sign_coding)
# remove sign strings which are not discrete, i.e. overlapping sign types 
# -> indicated by parentheses '()' 
coding.clean <- gsub("\\([^()]*\\)", "", coding.clean)
# remove sign strings which are not in linear order, i.e. those grouped in visual fields 
# -> indicated by square brackets '[]' 
coding.clean <- gsub("\\[(.*?)\\]", "", coding.clean)
# replace multiple underscores by single underscore
coding.clean <- gsub("( _ )+", " _ ", coding.clean)
# trim the underscores and white spaces from beginning and end of sequences
coding.clean <- trimws(coding.clean, which = "both", whitespace = "[ \\_]")

# add column to original data frame again
sign.data$coding_clean <- coding.clean
# again remove rows for which there is no sign coding (after cleaning)
sign.data <- sign.data[!sign.data$coding_clean == "", ]
```

```{r}
# use only lines with at least two characters
sign.data <- sign.data[nchar(sign.data$coding_clean) > 1, ]
```

Safe to file.
```{r}
write.csv(sign.data, 
          file = "~/Github/UpperPalCombinatorics/data/signBase/signBase_clean.csv",
          row.names = F)
```

