---
title: "Preprocessing TeDDi Sample"
author: "Chris Bentz"
date: "2/26/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables preprocessing of the TeDDi sample natural language text material. This is used as a sample to compare against the sign codings for proto-cuneiform and paleolithic sign sequences. Dumps of the TeDDi sample can be downloaded from https://drive.switch.ch/index.php/s/MJv7xFkzqlzFn0y. Note that the sample of TeDDi lines was here produced by using the follwing Perl commands in the directory where the lines.csv file is stored:

head -1 line.csv > line_sample.csv
perl -ne 'print if (rand() < .01)' line.csv >> line_sample.csv

This randomly samples 10% of the overall around 20 million text lines.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message=F}
library(dplyr)
```

# Load the TeDDi data
```{r}
# load data from TeDDi dumps (takes a couple of minutes)
line.data <- read.csv(
  "~/Github/UpperPalCombinatorics/data/TeDDi/line_sample.csv")
file.data <- read.csv("~/Data/TeDDi_dumps/file.csv")
# merge the files
teddi.data <- merge(line.data, file.data, by.x = "file_id", by.y = "id")
```

# Further Sampling
Sample a given number of lines randomly as a comparative sample.
```{r}
set.seed(15022018)
teddi.sample <-sample_n(teddi.data, 1000)
```

# Further Preprocessing
```{r}
# remove lines with less than 10 characters
teddi.clean <- teddi.sample[nchar(teddi.sample$text) > 10, ]
```

# Simple Stats
Some simple stats of the textual data here sampled from TeDDi.
```{r}
# lines of text
nrow(teddi.sample)
# unique ISO codes (i.e. languages)
unique(teddi.sample$iso639_3)
# writing systems
unique(teddi.sample$writing_system)
```

Write to file.
```{r}
write.csv(teddi.clean, 
          file = "~/Github/UpperPalCombinatorics/data/TeDDi/final_sample.csv", row.names = F)
```
