---
title: "Chinese Strokes versus Characters"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables replacing Chinese characters in vector of text lines with individual strokes represented by Latin characters (for easier further processing). The mapping from Chinese characters to strokes is taken from Wang et al. (2022), and can be downloaded from https://github.com/zjwang21/StrokeNet, the original file name is zh2letter.txt. 

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load Data
Load TeDDi subsample with all languages.
```{r}
teddi.data <- read.csv("~/Github/UpperPalCombinatorics/data/TeDDi/final_sample.csv")
# select Mandarin Chinese rows
cmn.data <- teddi.data[teddi.data$language_name_glotto == "Mandarin Chinese", ]
# remove Mandarin Chinese rows (to later append the cleaned rows)
teddi.data.nocmn <- teddi.data[teddi.data$language_name_glotto != "Mandarin Chinese", ]
```

# Transliterate Characters to Strokes

## Select Chinese lines
Select and clean lines with Chinese characters from the TeDDi subsample.
```{r}
# remove Latin script (English) contaminations
# the regex "[ -~]" matches ASCII characters
cmn.data$text <- gsub("[ -~]", '', cmn.data$text)
# remove punctuation (since this will give NAs when transliterating to strokes)
cmn.data$text <- gsub("[[:punct:]]", '', cmn.data$text)
# append cleaned rows to TeDDi data frame
teddi.data.modified <- rbind(teddi.data.nocmn, cmn.data) 
```

## Read file with strokes per character
Transliterate Chinese characters to strings of strokes represented by Latin characters.
```{r}
# load stroke transliteration list
stroke.file = "~/Github/UpperPalCombinatorics/data/TeDDi/zh2letter.txt"
stroke.df <- read.table("~/Github/UpperPalCombinatorics/data/TeDDi/zh2letter.txt", sep = " ")
# add column names
colnames(stroke.df) <- c("chars","strokes")
# give beginning of signlist
head(stroke.df)
```

## Replace Chinese characters by strokes
```{r}
# split into individual characters (list by text line)
cmn.list <- strsplit(cmn.data$text, split = "")
head(cmn.list)
# replace characters with stroke transliterations for each string in list
for (i in 1: length(cmn.list)){
  cmn.list[[i]] <- paste(with(stroke.df, strokes[match(cmn.list[[i]], chars)]), collapse = " ")
}
head(cmn.list)
# unlist and replace original text column in cmn dataframe
cmn.data$text <- unlist(cmn.list)
# add comment on stroke transliteration in column "comments"
cmn.data$comments <- rep("chars transliterated to strokes", nrow(cmn.data))
# add cmn data with strokes to overall data frame
teddi.data.modified <- rbind(teddi.data.modified, cmn.data) 
```

Write to file.
```{r}
write.csv(teddi.data.modified, 
          file = "~/Github/UpperPalCombinatorics/data/TeDDi/final_sample_strokes.csv", row.names = F)
```

# Literature References
Wang, Z., Liu, X., & Zhang, M. (2022). Breaking the representation bottleneck of Chinese characters: Neural machine translation with stroke sequence modeling. arXiv preprint arXiv:2211.12781.
