---
title: "Randomization"
author: ''
date: "28/03/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file enables concatenation and randomization of sign strings which occur on different parts of a paleolithic object (indicated by underscore `_'). In the case of paleolithic sequences, we do not know the order of these respective strings (if any), hence, if we want to analyse them as one combined sequence, they need to be randomized. Note that this is not applicable to modern day written languages, as the order of strings of characters (e.g. words) is meaningful. Randomization of strings would artificially introduce randomness where order is intended. Arguably, the same holds for proto-cuneiform writing. There are particular orders with which proto-cuneiform signs were imprinted onto a tablet, i.e. a kind of "syntax", cf. the following statement in Englund (1998), p. 79:

"Number sign sequences within discrete notations are, as might be expected, very rigid and follow a definite numerical `syntax’. Within text entries, moreover, the position of numerical notations relative to ideographic notations is fairly rigid."

Again, randomization would destroy this syntax and artificially introduce randomness where order was intended by the scribes.  

## References
Englund (1998). Texts from the Late Uruk period. In Bauer et al. (Eds.), Späturuk-Zeit und Frühdynastische Zeit.

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). 
```{r, message = F}
#library(stringr)
```

# Load Data
```{r}
sign.data <- read.csv("~/Github/UpperPalCombinatorics/data/signBase/signBase_clean.csv")
head(sign.data)
```

# Randomize
Randomize sign sequences per object.
```{r}
# set seed to get reproducible result
set.seed(09012020)
# initialize empty vector
coding.random <- c()
# run loop to randomize, and add new column to data frame
for (i in 1:nrow(sign.data)){
  # use single coding per row
  coding.local <- sign.data[i, ]$coding_clean
  # split on underscores surrounded by white spaces
  coding.split <- unlist(strsplit(as.character(coding.local), split = " _ "))
  # randomize order
  coding.random.local <- sample(coding.split)
  # concatenate with underscore and white space again
  coding.random.local <- paste(coding.random.local, collapse = " _ ")
  # add to vector
  coding.random <- c(coding.random, coding.random.local)
}

# add column with randomized order to data frame
sign.data$coding_random <- coding.random
```

Safe to file.
```{r}
write.csv(sign.data, "~/Github/UpperPalCombinatorics/data/signBase/signBase_randomized.csv", row.names = F)
```


