---
title: "Estimations of Quantitative Features for TeDDi Sample"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file contains the code for estimating the feature values (TTR, unigram entropy, entropy rate, repetition rate) per character sequence in TeDDi sample lines. These features where tested on a wider sample of linguistic and non-linguistic sign sequences in Bentz (2023). The repetition measure was tested also in Sproat (2014). 

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(entropy)
library(gsubfn)
#library(devtools)
#install_github("dimalik/Hrate")
library(Hrate)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# External functions
```{r sourceCode, echo = TRUE}
# repetition rate function rrate()
source("~/Github/UpperPalCombinatorics/code/functions/repRate_function.R")
```

# Load files
Load the file with cleaned TeDDi sample lines.
```{r}
teddi.clean <- read.csv("~/Github/UpperPalCombinatorics/data/TeDDi/final_sample.csv")
```

# Feature Estimations
```{r}
# set counter
counter = 0
# initialize dataframe to append results to
estimations.df <- data.frame(id = numeric(0), text = character(0), filename = character(0),
                             language_name_glotto = character(0), iso639_3 = character(0),
                             mode = character(0), genre_broad = character(0), 
                             writing_system = character(0), comments = character(0),
                             num.tokens = numeric(0), num.types = numeric (0),  
                             huni.chars = numeric (0), hrate.chars = numeric(0), 
                             ttr.chars = numeric(0), rm.chars = numeric(0)) 

# start time
start_time <- Sys.time()
for (i in 1:nrow(teddi.clean)){
  try({ # if the processing fails for a certain row, there will be no output for this row,
  # but the try() function allows the loop to keep running  
  
  # collect relevant meta info from row
  id <- teddi.clean$id[i]
  coding.sample <- as.character(teddi.clean$text[i])
  text <- coding.sample
  filename <- teddi.clean$filename[i]
  language_name_glotto <- teddi.clean$language_name_glotto[i]
  iso639_3 <- teddi.clean$iso639_3[i]
  mode <- teddi.clean$mode[i]
  genre_broad <- teddi.clean$genre_broad[i]
  writing_system <- teddi.clean$writing_system[i]
  comments <- teddi.clean$comments[i]

  # split strings into UTF-8 characters
  chars <- unlist(strsplit(coding.sample, ""))
  
  # get number of sign tokens and types
  num.tokens <- length(chars)
  num.types <- length(unique(chars))
  
  # preprocess text with bag-of-words normalization
  # note: this is an important step! Without it, entropy rate estimation will fail
  # to yield realistic results due to issues with processing special characters.
  chars <- PreprocessText(chars)
  
  # unigram entropy estimation
  # calculate unigram entropy for characters
  chars.df <- as.data.frame(table(chars))
  # print(chars.df)
  huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
  
  # entropy rate estimation
  # the values chosen for max.length and every.word will crucially 
  # impact processing time. In case of "max.length = NULL" the length is n units
  # here use the try() function since entropy rate estimation can fail when particular
  # special characters are in the strings
  hrate.chars <- try(get.estimate(text = chars, every.word = 1, max.length = NULL))
  if (class(hrate.chars) == "numeric") {
    hrate.chars <- hrate.chars
  } else {
    hrate.chars = "NA"
  }
  
  # calculate type-token ratio (TTR)
  ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
  
  # calculate repetition rate
  rm.chars <- rrate(chars, method = "lsmo")
  
  # append results to dataframe
  local.df <- data.frame(id, text, filename, language_name_glotto,
                         iso639_3, mode, genre_broad, 
                         writing_system, comments,
                         num.tokens, num.types, huni.chars, hrate.chars,
                         ttr.chars, rm.chars)
  estimations.df <- rbind(estimations.df, local.df)
}) # end of try()
# counter
counter <- counter + 1
# print(counter)
}
end_time <- Sys.time()
end_time - start_time
```

Give head of estimation data frame (selected columns) for visual checking of the output data structure.
```{r}
head(estimations.df)
```

Safe outputs to file.
```{r}
write.csv(estimations.df, "~/Github/UpperPalCombinatorics/results/features/teddi_features.csv", 
          row.names = F)
```

# Package References
```{r, echo = F}
print(citation("Hrate"), style = "text")
print(citation("stringr"), style = "text")
print(citation("gsubfn"), style = "text")
print(citation("entropy"), style = "text")
```

# Literature References
Bentz, C. (2023). The Zipfian Challenge: Learning the statistical fingerprint of natural languages. In Proceedings of the 27th Conference on Computational Natural Language Learning (CoNLL), pp. 27-37.

Sproat, R. (2014). A statistical comparison of written language and nonlinguistic symbol systems. Language, 457-481.
