---
title: "Estimations of Quantitative Features for Cuneiform Tablets"
author: ""
date: "27/02/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
This file contains the code for estimating the feature values (TTR, unigram entropy, entropy rate, repetition rate) per sign sequence on cuneiform tablets from CDLI.

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(entropy)
library(gsubfn)
#library(devtools)
#install_github("dimalik/Hrate")
library(Hrate)
```

# Load files
Load the files with the cleaned sequences.
```{r}
# choose period (UrukIII, UrukIV, UrukV)
period <- "UrukIII"

# cuneiform data from CDLI
cunei.data <- read.csv(paste("~/Github/UpperPalCombinatorics/data/CDLI/", 
                      paste(period, "/signs_clean.csv", sep = ""), 
                      sep =""))
# remove rows with empty sign codings
cunei.data <- cunei.data[-which(cunei.data$signs.orig.vec == ""), ]
```

# Cuneiform Estimations
Estimations of quantitative features for the cuneiform data from CDLI.
```{r}
# set counter
counter = 0
# initialize dataframe to append results to
estimations.df <- data.frame(tablet.id = character(0), num.char = numeric(0),
                             huni.chars = numeric (0), hrate.chars = numeric(0),
                             ttr.chars = numeric(0), rm.chars = numeric(0)) 

# start time
start_time <- Sys.time()
for (i in 1:nrow(cunei.data)){
  try({ # if the processing fails for a certain row, there will be no output for this row,
  # but the try() function allows the loop to keep running  
  
  # collect info from row
  tablet.id <- cunei.data$tablet.id[i]
  coding.sample <- as.character(cunei.data$signs.expand.vec[i])

  # split strings into cuneiform sign types
  chars <- unlist(strsplit(coding.sample, " "))
  
  # get overall number of signs on tablet
  num.char <- length(chars)
  
  # preprocess the signs with bag-of-words normalization
  # note: this is an important step! Without it, entropy rate estimation will fail
  # to yield realistic results due to issues with processing special characters.
  chars <- PreprocessText(chars)
  
  # unigram entropy estimation
  # calculate unigram entropy for characters (note that this is the same
  # for random ordering of chracter strings)
  chars.df <- as.data.frame(table(chars))
  # print(chars.df)
  huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
  
  # entropy rate estimation
  # the values chosen for max.length and every.word will crucially 
  # impact processing time. In case of "max.length = NULL" the length is n units
  # here use the try() function since entropy rate estimation can fail when particular
  # special characters are in the strings
  hrate.chars <- try(get.estimate(text = chars, every.word = 1, max.length = NULL))
  if (class(hrate.chars) == "numeric") {
    hrate.chars <- hrate.chars
  } else {
    hrate.chars = "NA"
  }
  # calculate type-token ratio (ttr) 
  # note that this is the same for random ordering of strings
  ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
  
  # calculate repetition measure according to Sproat (2014)
  # note that this is the same for random ordering of strings
  # the maximum number of adjacent repetitions "R" possible is the sum of frequency counts minus 1. 
  R <- sum(chars.df$Freq)-1 
  # calculate the number of adjacent repetitions "r"
  r = 0
  if (length(chars) > 1){
    for (i in 1:(length(chars)-1)){
      if (chars[i] == chars[i+1]){
        r = r + 1
      } else {
        r = r + 0
      }
      }
    # calculate the repetition measure
    rm.chars <- r/R
  } else {
    rm.chars <- "NA"  
  }
  
  # append results to dataframe
  local.df <- data.frame(tablet.id, num.char, huni.chars, hrate.chars,
                         ttr.chars, rm.chars)
  estimations.df <- rbind(estimations.df, local.df)
}) # end of try()
# counter
counter <- counter + 1
# print(counter)
}
end_time <- Sys.time()
end_time - start_time
#estimations.df
```

Merge estimations with original file
```{r}
cunei.data.est <- merge(cunei.data, estimations.df, by = "tablet.id")
```

Safe outputs to file.
```{r}
write.csv(cunei.data.est, paste("~/Github/UpperPalCombinatorics/results/cuneiform_", 
                      paste(period, "_features.csv", sep = ""), 
                      sep =""), row.names = F)
```
