---
title: "Estimations of Quantitative Features for Cuneiform Tablets"
author: "Chris Bentz"
date: "06.03.2025"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This file contains the code for estimating the feature values (TTR, unigram entropy, entropy rate, repetition rate) per sign sequence on proto-cuneiform tablets from CDLI. These features where tested on a wider sample of linguistic and non-linguistic sign sequences in Bentz (2023). The repetition measure was tested also in Sproat (2014). 

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(entropy)
library(gsubfn)
#library(devtools)
#install_github("dimalik/Hrate")
library(Hrate)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# External functions
```{r sourceCode, echo = TRUE}
# repetition rate function rrate()
source("~/Github/PaleoSigns/code/functions/repRate_function.R")
```

# Load files
Load the files with the cleaned sequences. A period needs to be chosen here. Hence, the code needs to be run for periods separately.
```{r}
# choose period (UrukIII, UrukIV, UrukV)
period <- "UrukV"

# cuneiform data from CDLI
cunei.data <- read.csv(paste("~/Github/PaleoSigns/data/CDLI/", 
                      paste(period, "/signs_clean.csv", sep = ""), 
                      sep =""))
# head(cunei.data)
```

# Cuneiform Estimations
Estimations of quantitative features for the cuneiform data from CDLI.
```{r}
# set counter
counter = 0
# initialize dataframe to append results to
estimations.df <- data.frame(tablet.id = character(0), 
                             num.tokens = numeric(0),
                             num.types = numeric(0),
                             huni.chars = numeric (0), 
                             hrate.chars = numeric(0), 
                             ttr.chars = numeric(0), 
                             rm.chars = numeric(0)) 

# start time
start_time <- Sys.time()
for (i in 1:nrow(cunei.data)){
  try({ # if the processing fails for a certain row, there will be no output for this row,
  # but the try() function allows the loop to keep running  
  
  # collect info from row
  tablet.id <- cunei.data$tablet.id[i]
  coding.sample <- as.character(cunei.data$signs.expand.vec[i])

  # split sequences per tablet into cuneiform signs
  # this is here done by using white spaces
  # note: sign transliterations consist of multiple characters,
  # e.g. NO1, N14, SZE~a, etc.
  chars <- unlist(strsplit(coding.sample, " "))
  
  # get number of tokens and types, i.e. "characters" are proto-cuneiform signs in this case
  num.tokens <- length(chars)
  num.types <- length(unique(chars))
  
  # preprocess the signs with bag-of-words normalization
  # note: this is an important step! Without it, entropy rate estimation will fail
  # to yield realistic results due to issues with processing special characters.
  chars <- PreprocessText(chars)
  
  # unigram entropy estimation
  # calculate unigram entropy for characters
  chars.df <- as.data.frame(table(chars))
  # print(chars.df)
  huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
  
  # entropy rate estimation
  # the values chosen for max.length and every.word will crucially impact processing time 
  # in case of "max.length = NULL" the length the total length of the sequence (n)
  # use the try() function since entropy rate estimation can fail when particular
  # special characters are in the strings
  hrate.chars <- try(get.estimate(text = chars, every.word = 1, max.length = NULL))
  if (class(hrate.chars) == "numeric") {
    hrate.chars <- hrate.chars
  } else {
    hrate.chars = "NA"
  }
  # calculate type-token ratio (TTR) 
  ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
  
 # calculate repetition rate
  rm.chars <- rrate(chars, method = "lsmo")
  
  # append results to dataframe
  local.df <- data.frame(tablet.id, num.tokens, num.types, huni.chars, hrate.chars,
                         ttr.chars, rm.chars)
  estimations.df <- rbind(estimations.df, local.df)
}) # end of try()
# counter
counter <- counter + 1
# print(counter)
}
end_time <- Sys.time()
end_time - start_time
```

Merge estimations with original file
```{r}
cunei.data.est <- merge(cunei.data, estimations.df, by = "tablet.id")
```

Give head of estimation data frame (selected columns) for visual checking of the output data structure.
```{r}
# head(cunei.data.est[c("tablet.id", "signs.orig.vec", "num.tokens", "num.types", 
#                      "huni.chars",  "hrate.chars", "ttr.chars", "rm.chars")])
```

Safe outputs to file.
```{r}
write.csv(cunei.data.est, paste("~/Github/PaleoSigns/results/features/cuneiform_", 
                      paste(period, "_features.csv", sep = ""), 
                      sep =""), row.names = F)
```

# Package References
```{r, echo = F}
print(citation("Hrate"), style = "text")
print(citation("stringr"), style = "text")
print(citation("gsubfn"), style = "text")
print(citation("entropy"), style = "text")
```

# Literature References
Bentz, C. (2023). The Zipfian Challenge: Learning the statistical fingerprint of natural languages. In Proceedings of the 27th Conference on Computational Natural Language Learning (CoNLL), pp. 27-37.

Sproat, R. (2014). A statistical comparison of written language and nonlinguistic symbol systems. Language, 457-481.

