---
title: "Estimations Combined"
author: ""
date: "2/28/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This code loads and combines the different feature value estimations (natural language writing, paleolithic signs, and proto-cuneiform) into one csv file.

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
#library(stringr)
library(dplyr)
```

# Load Data
Load csv files with quantitative feature values for sign sequences. 

## SIGNBASE
```{r}
# load signbase estimations for strings on objects
sign.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/signBase_features.csv")
# select Aurignacian objects
sign.data <- sign.data[sign.data$archaeological_unit == "Aurignacien", ] 

# select columns for direct comparison with TeDDi sample
sign.data.selected <- select(sign.data, c('archaeological_unit', 'coding_clean', 
                                          'num.char', 'huni.chars', 'hrate.chars', 
                                 'ttr.chars', 'rm.chars'))
# change column names
colnames(sign.data.selected) <- c('subcorpus', 'sequence', 'num.char','huni.chars', 'hrate.chars', 
                         'ttr.chars', 'rm.chars')
```

## CDLI
```{r}
# read feature values for proto-cuneiform from CDLI
# Uruk V
cuneiV.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/cuneiform_UrukV_features.csv")

# add column with subcorpus information
cuneiV.data$subcorpus <- rep("Cuneiform (Uruk V)", nrow(cuneiV.data))

# select columns for direct comparison with other samples
cuneiV.data.selected <- select(cuneiV.data, c('subcorpus', 'signs.expand.vec', 
                                              'num.char', 'huni.chars', 'hrate.chars', 
                                              'ttr.chars', 'rm.chars'))
# rename column with sign sequences
names(cuneiV.data.selected)[names(cuneiV.data.selected) == "signs.expand.vec"] <- "sequence"

# Uruk IV
cuneiIV.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/cuneiform_UrukIV_features.csv")

# add column with subcorpus information
cuneiIV.data$subcorpus <- rep("Cuneiform (Uruk IV)", nrow(cuneiIV.data))

# select columns for direct comparison with other sample
cuneiIV.data.selected <- select(cuneiIV.data, c('subcorpus', 'signs.expand.vec', 
                                              'num.char', 'huni.chars', 'hrate.chars', 
                                              'ttr.chars', 'rm.chars'))
# rename column with sign sequences
names(cuneiIV.data.selected)[names(cuneiIV.data.selected) == "signs.expand.vec"] <- "sequence"

# Uruk III
cuneiIII.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/cuneiform_UrukIII_features.csv")

# add column with subcorpus information
cuneiIII.data$subcorpus <- rep("Cuneiform (Uruk III)", nrow(cuneiIII.data))

# select columns for direct comparison with other sample
cuneiIII.data.selected <- select(cuneiIII.data, c('subcorpus', 'signs.expand.vec', 
                                              'num.char', 'huni.chars', 'hrate.chars', 
                                              'ttr.chars', 'rm.chars'))
# rename column with sign sequences
names(cuneiIII.data.selected)[names(cuneiIII.data.selected) == "signs.expand.vec"] <- "sequence"
```

## TeDDI
```{r}
# read feature values for character sequences of the TeDDi sample
teddi.data <- read.csv("~/Github/UpperPalCombinatorics/results/features/teddi_features.csv")
# add column with subcorpus information
teddi.data$subcorpus <- rep("TeDDi", nrow(teddi.data))
# select columns
teddi.selected <- select(teddi.data, c('subcorpus','text', 'num.char','huni.chars', 'hrate.chars', 
                           'ttr.chars', 'rm.chars'))
# rename column with sign sequences
names(teddi.selected)[names(teddi.selected) == "text"] <- "sequence"
```

# Combine
Combine the different data frames.
```{r}
estimations.df <- rbind(sign.data.selected, cuneiV.data.selected, 
                        cuneiIV.data.selected, cuneiIII.data.selected,
                        teddi.selected)
head(estimations.df)
```

# Simple Stats
```{r}
nrow(estimations.df) # number of sequences with feature estimations
unique(estimations.df$subcorpus) # names of subcorpora
nrow(estimations.df[estimations.df$subcorpus=="Aurignacien",]) # number of Aurignacien sequences
nrow(estimations.df[estimations.df$subcorpus=="TeDDi",]) # number of TeDDi (modern writing) sequences
nrow(estimations.df[estimations.df$subcorpus=="Cuneiform (Uruk V)",]) 
nrow(estimations.df[estimations.df$subcorpus=="Cuneiform (Uruk IV)",]) 
nrow(estimations.df[estimations.df$subcorpus=="Cuneiform (Uruk III)",]) 
```

Write to file
```{r}
write.csv(estimations.df, "~/Github/UpperPalCombinatorics/results/features/features_combined.csv", row.names = F)
```